---
title: "Changing in Equity and Leadership in Japan–Vietnam Scientific Collaboration"
subtitle: "A Bibliometric Assessment — Analysis"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    df-print: paged
    fig-width: 9
    fig-height: 5
execute:
  echo: true
  warning: false
  message: false
  cache: false
  freeze: auto
---

# Setup


```{r}
#| label: setup
#| include: false

library(tidyverse)
library(readr)
library(scales)
library(lubridate)

set.seed(20260105)

options(
  dplyr.summarise.inform = FALSE,
  readr.show_col_types = FALSE,
  stringsAsFactors = FALSE
)

DATA_DIR <- "data"
OUT_DIR  <- "outputs"
FIG_DIR  <- file.path(OUT_DIR, "figures")
TAB_DIR  <- file.path(OUT_DIR, "tables")

dir.create(OUT_DIR, showWarnings = FALSE, recursive = TRUE)
dir.create(FIG_DIR, showWarnings = FALSE, recursive = TRUE)
dir.create(TAB_DIR, showWarnings = FALSE, recursive = TRUE)

theme_set(
  theme_minimal(base_size = 12, base_family = "sans") +
    theme(
      plot.title.position   = "plot",
      plot.caption.position = "plot",
      legend.position       = "bottom",
      legend.title          = element_text(size = 10),
      legend.text           = element_text(size = 10),
      axis.title            = element_text(size = 11),
      axis.text             = element_text(size = 10),
      strip.text            = element_text(size = 11),
      panel.grid.minor      = element_blank()
    )
)

ggplot2::update_geom_defaults("line",   list(linewidth = 0.9))
ggplot2::update_geom_defaults("smooth", list(linewidth = 0.9))
ggplot2::update_geom_defaults("col",    list(linewidth = 0))
ggplot2::update_geom_defaults("area",   list(linewidth = 0))
```

# Data and samples

```{r}
#| label: load-data

DATA_FILE <- file.path(DATA_DIR, "updated_data.csv")

dat <- read_csv(DATA_FILE) |>
  mutate(
    year = as.integer(year),
    coop = factor(coop, levels = c("Bilateral", "Multilateral", "Hyperauthored")),
    quartile = factor(quartile, levels = c("Q1", "Q2", "Q3", "Q4"), ordered = TRUE),
    oa = as.integer(oa),
    fund = as.integer(fund),
    across(c(LS, SS, PS, HS), as.integer)
  ) |>
  filter(!is.na(coop))

dat_core <- dat |> filter(coop != "Hyperauthored")

dat_bi <- dat |> filter(coop == "Bilateral")
```


```{r}
#| label: tbl-profile

fmt_iqr <- function(x) {
  qs <- quantile(x, probs = c(0.25, 0.50, 0.75), na.rm = TRUE, type = 2)
  paste0(scales::comma(qs[[2]]), " (", scales::comma(qs[[1]]), "–", scales::comma(qs[[3]]), ")")
}

fmt_n_pct <- function(n, d) {
  if (is.na(d) || d == 0) return(paste0(scales::comma(n), " (NA)"))
  paste0(scales::comma(n), " (", sprintf("%.1f", 100 * n / d), "%)")
}

mode_stats <- bind_rows(
  dat |> mutate(mode = as.character(coop)),
  dat |> mutate(mode = "Total")
  ) |>
  group_by(mode) |>
  summarise(
    n = n(),
    n_auth_iqr = fmt_iqr(n_authors),
    n_ctry_iqr = fmt_iqr(n_countries),
    oa_n = sum(oa == 1, na.rm = TRUE),
    fund_n = sum(fund == 1, na.rm = TRUE),
    q1_n = sum(quartile == "Q1", na.rm = TRUE),
    q2_n = sum(quartile == "Q2", na.rm = TRUE),
    q3_n = sum(quartile == "Q3", na.rm = TRUE),
    q4_n = sum(quartile == "Q4", na.rm = TRUE),
    q_na_n = sum(is.na(quartile)),
    LS_n = sum(LS == 1, na.rm = TRUE),
    SS_n = sum(SS == 1, na.rm = TRUE),
    PS_n = sum(PS == 1, na.rm = TRUE),
    HS_n = sum(HS == 1, na.rm = TRUE),
  .groups = "drop"
  )

make_row <- function(metric, values) {
  tibble(metric = metric) |>
  bind_cols(as_tibble_row(setNames(values, mode_stats$mode)))
}

tbl1 <- bind_rows(
  make_row("N papers", mode_stats$n |> scales::comma()),
  make_row("median n_authors (IQR)", mode_stats$n_auth_iqr),
  make_row("median n_countries (IQR)", mode_stats$n_ctry_iqr),
  make_row("OA n (%)", map2_chr(mode_stats$oa_n, mode_stats$n, fmt_n_pct)),
  make_row("funded n (%)", map2_chr(mode_stats$fund_n, mode_stats$n, fmt_n_pct)),
  make_row("Q1 n (% within mode)", map2_chr(mode_stats$q1_n, mode_stats$n, fmt_n_pct)),
  make_row("Q2 n (% within mode)", map2_chr(mode_stats$q2_n, mode_stats$n, fmt_n_pct)),
  make_row("Q3 n (% within mode)", map2_chr(mode_stats$q3_n, mode_stats$n, fmt_n_pct)),
  make_row("Q4 n (% within mode)", map2_chr(mode_stats$q4_n, mode_stats$n, fmt_n_pct)),
  make_row("Quartile missing n (% within mode)", map2_chr(mode_stats$q_na_n, mode_stats$n, fmt_n_pct)),
  make_row("LS n (% within mode)", map2_chr(mode_stats$LS_n, mode_stats$n, fmt_n_pct)),
  make_row("SS n (% within mode)", map2_chr(mode_stats$SS_n, mode_stats$n, fmt_n_pct)),
  make_row("PS n (% within mode)", map2_chr(mode_stats$PS_n, mode_stats$n, fmt_n_pct)),
  make_row("HS n (% within mode)", map2_chr(mode_stats$HS_n, mode_stats$n, fmt_n_pct))
  ) |>
  select(metric, Bilateral, Multilateral, Hyperauthored, Total)

tbl1 |> knitr::kable(align = c("l", "r", "r", "r", "r"))
```

# RQ1 — Trajectory

```{r}
#| label: fig-volume

d <- dat |> filter(year >= 2000)

vol_year <- d |>
  count(year, name = "n_total") |>
  arrange(year) |>
  mutate(growth = n_total / lag(n_total) - 1)

mode_year <- d |>
  count(year, coop, name = "n_mode") |>
  left_join(vol_year, by = "year")

min_g <- suppressWarnings(min(vol_year$growth, na.rm = TRUE))
max_g <- suppressWarnings(max(vol_year$growth, na.rm = TRUE))
max_n <- max(vol_year$n_total, na.rm = TRUE)

vol_year <- vol_year |>
  mutate(growth_scaled = (growth - min_g) / (max_g - min_g) * max_n)

ggplot() +
  geom_col(
    data = mode_year, 
    aes(x = year, y = n_mode, fill = coop)
  ) +
  geom_line(
    data = vol_year,
    aes(x = year, y = growth_scaled, group = 1),
    linewidth = 0.5
  ) +
  geom_vline(xintercept = 2014, linetype = "dotted") +
  geom_vline(xintercept = 2023, linetype = "dotted") +
  scale_y_continuous(
    name = "Publications (count)",
    sec.axis = sec_axis(
      ~ . / max_n * (max_g - min_g) + min_g,
      name = "Annual growth rate",
      labels = scales::percent_format(accuracy = 1)
    )
  ) +
  labs(
    x = NULL, 
    fill = "Collaboration mode"
  )
```


```{r}
#| label: fig-fields

dat |>
  filter(year >= 2000) |>
  group_by(year) |>
  summarise(
    LS = mean(LS == 1, na.rm = TRUE),
    SS = mean(SS == 1, na.rm = TRUE),
    PS = mean(PS == 1, na.rm = TRUE),
    HS = mean(HS == 1, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_longer(
    cols = c(LS, SS, PS, HS), 
    names_to = "field", 
    values_to = "share"
  ) |>
  mutate(field = factor(field, levels = c("LS", "SS", "PS", "HS"))) |>
  ggplot(aes(x = year, y = share, color = field)) +
  geom_line() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = NULL, 
    y = "Share of publications", 
    color = NULL
  )
```


```{r}
#| label: fig-fields-stacked-normalized

dat |>
  filter(year >= 2000) |>
  group_by(year) |>
  summarise(
    LS = sum(LS == 1, na.rm = TRUE),
    SS = sum(SS == 1, na.rm = TRUE),
    PS = sum(PS == 1, na.rm = TRUE),
    HS = sum(HS == 1, na.rm = TRUE),
    .groups = "drop"
  ) |>
  # Normalize so the sum of LS, SS, PS, HS for each year equals 1
  mutate(total_hits = LS + SS + PS + HS) |>
  mutate(across(c(LS, SS, PS, HS), ~ . / total_hits)) |>
  select(-total_hits) |>
  pivot_longer(
    cols = c(LS, SS, PS, HS), 
    names_to = "field", 
    values_to = "share"
  ) |>
  mutate(field = factor(field, levels = c("LS", "PS", "HS", "SS"))) |>
  ggplot(aes(x = year, y = share, fill = field)) +
  geom_area(alpha = 0.8) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1), 
    expand = c(0, 0)
  ) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    x = NULL, 
    y = "Relative Research Effort", 
    fill = "Field"
  )
```


```{r}
#| label: fig-quality

dat |>
  filter(year >= 2000, !is.na(quartile)) |>
  count(year, quartile, name = "n") |>
  group_by(year) |>
  mutate(share = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(x = year, y = share, fill = quartile)) +
  geom_area(alpha = 0.7) +
  coord_cartesian(ylim = c(0, 1)) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1), 
    breaks = seq(0, 1, 0.2),
    expand = c(0, 0)
  ) +
  labs(
    x = NULL, 
    y = "Share of publications", 
    fill = NULL
  )
```


```{r}
#| label: fig-oa

dat |>
  filter(year >= 2000) |>
  mutate(
    oa_status = factor(if_else(oa == 1, "OA", "Non-OA"), levels = c("OA", "Non-OA"))
  ) |>
  count(year, oa_status, name = "n") |>
  group_by(year) |>
  mutate(share = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(x = year, y = share, fill = oa_status)) +
  geom_area(alpha = 0.85) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "black") +
  coord_cartesian(ylim = c(0, 1)) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1), 
    breaks = seq(0, 1, 0.2),
    expand = c(0, 0)
  ) +
  labs(
    x = NULL, 
    y = "Share of publications", 
    fill = NULL
  )
```

# RQ2


```{r}
#| label: fig-equity

dat <- dat |> mutate(parity_gap = abs(fc_vn - fc_jp))

dat |>
  filter(year >= 2000) |>
  group_by(year) |>
  summarise(
    fc_vn = mean(fc_vn, na.rm = TRUE),
    fc_jp = mean(fc_jp, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_longer(
    cols = c(fc_vn, fc_jp), 
    names_to = "series", 
    values_to = "value"
  ) |>
  mutate(
    series = recode(
      series,
      fc_vn = "Vietnam (fractional)",
      fc_jp = "Japan (fractional)"
    ),
    series = factor(
      series, 
      levels = c("Vietnam (fractional)", "Japan (fractional)")
    )
  ) |>
  {
    \(data) {
      lim <- range(data$value, na.rm = TRUE)
      pad <- max(0.02, 0.08 * diff(lim))
      
      ggplot(data, aes(x = year, y = value, color = series)) +
        geom_line() +
        coord_cartesian(ylim = c(lim[1] - pad, lim[2] + pad)) +
        scale_y_continuous(labels = percent_format(accuracy = 1)) +
        # MODIFIED: Use manual colors to match Figure 2
        scale_color_manual(values = c(
          "Vietnam (fractional)" = "#1B9E77", # Green
          "Japan (fractional)" = "#D95F02"    # Orange
        )) +
        labs(
          x = NULL, 
          y = "Mean fractional share", 
          color = NULL
        )
    }
  }()
```


```{r}
#| label: fig-leadership

lead_year <- bind_rows(
  dat |>
    filter(year >= 2000) |>
    group_by(year) |>
    summarise(value = mean(fa_vn, na.rm = TRUE), .groups = "drop") |>
    mutate(series = "Vietnam — First author"),
  dat |>
    filter(year >= 2000) |>
    group_by(year) |>
    summarise(value = mean(fa_jp, na.rm = TRUE), .groups = "drop") |>
    mutate(series = "Japan — First author"),
  dat |>
    filter(year >= 2000, !is.na(ca_vn)) |>
    group_by(year) |>
    summarise(value = mean(ca_vn, na.rm = TRUE), .groups = "drop") |>
    mutate(series = "Vietnam — Corresponding author"),
  dat |>
    filter(year >= 2000, !is.na(ca_jp)) |>
    group_by(year) |>
    summarise(value = mean(ca_jp, na.rm = TRUE), .groups = "drop") |>
    mutate(series = "Japan — Corresponding author")
) |>
  mutate(
    series = factor(
      series,
      levels = c(
        "Vietnam — First author",
        "Japan — First author",
        "Vietnam — Corresponding author",
        "Japan — Corresponding author"
      )
    )
  )

lead_year |>
  {
    \(data) {
      lim <- range(data$value, na.rm = TRUE)
      pad <- max(0.02, 0.08 * diff(lim))
      
      ggplot(data, aes(x = year, y = value, color = series, linetype = series)) +
        geom_line() +
        coord_cartesian(ylim = c(lim[1] - pad, lim[2] + pad)) +
        scale_y_continuous(labels = percent_format(accuracy = 1)) +
        # MODIFIED: Assign same color to same country, regardless of author type
        scale_color_manual(values = c(
          "Vietnam — First author" = "#1B9E77",         # Green
          "Vietnam — Corresponding author" = "#1B9E77", # Green
          "Japan — First author" = "#D95F02",           # Orange
          "Japan — Corresponding author" = "#D95F02"    # Orange
        )) +
        # Keep linetypes to distinguish author role
        scale_linetype_manual(values = c("dashed", "dashed", "solid", "solid")) +
        labs(
          x = NULL, 
          y = "Leadership share", 
          color = NULL, 
          linetype = NULL
        )
    }
  }()
```

# RQ3

This chunk constructs a time- and field-aware citation impact normalization that preserves the multi-label nature of Scopus macro-field tagging (HS, LS, SS, PS). Because many papers are assigned to multiple macro fields, forcing a single-field classification would be arbitrary and could distort comparisons. We therefore compute year-specific field baselines and then normalize each paper relative to the expected citation level implied by its own field portfolio.

Concretely, we first reshape the dataset into a long format to estimate, for each year × macro field, the average citation count (`mu_cited`) and the average log-citation (`mu_log1p = mean(log1p(cited))`). Next, for each paper, we calculate the number of macro fields it is tagged in (`k_field`) and assign equal fractional weights (1/`k_field`) across its tagged fields. Using these weights, we derive an expected baseline for each paper as the weighted average of the relevant year × field baselines. Finally, we construct two normalized impact measures: (i) a ratio-based index, `impact_norm = cited / impact_mu`, and (ii) a log-centered index, `impact_norm_log = log1p(cited) − impact_mu_log`. The log-centered form is typically more stable in citation data because it reduces sensitivity to extreme outliers and heavy right tails, while still retaining the interpretation of relative performance against a year- and field-adjusted benchmark.

Let paper (i) be published in year (y_i). Let the set of macro fields be (F={\text{HS},\text{LS},\text{SS},\text{PS}}). Define an indicator (I_{if}\in{0,1}) that paper (i) is tagged in field (f).

### Step 1. Year–field baselines

[
\mu^{(c)}*{f,y}=\mathbb{E}!\left[C \mid f,y\right]
]
[
\mu^{(\ell)}*{f,y}=\mathbb{E}!\left[\log(1+C)\mid f,y\right]
]
where (C) is the citation count.

### Step 2. Fractional weights for multi-field papers

[
k_i=\sum_{f\in F} I_{if}
\qquad\text{and}\qquad
w_{if}=\frac{I_{if}}{k_i}
]

### Step 3. Paper-specific expected baselines (mixture over fields)

[
\mu^{(c)}*{i}=\sum*{f\in F} w_{if},\mu^{(c)}*{f,y_i}
]
[
\mu^{(\ell)}*{i}=\sum_{f\in F} w_{if},\mu^{(\ell)}_{f,y_i}
]

### Step 4. Normalized impact measures

**Ratio normalization**
[
\text{impact_norm}*i=\frac{C_i}{\mu^{(c)}*{i}}
]

**Log-centered normalization**
[
\text{impact_norm_log}*i=\log(1+C_i)-\mu^{(\ell)}*{i}
]

These are exactly the quantities computed in the chunk: the paper’s expected baseline is the weighted average of year×field means, with equal fractional weight across all macro fields assigned to that paper.


```{r}
#| label: impact-norm-func

macro_fields <- c("HS", "LS", "SS", "PS")

make_impact_norm <- function(df) {
  field_year_mu <- df |>
    pivot_longer(all_of(macro_fields), names_to = "field", values_to = "flag") |>
    filter(flag == 1) |>
    group_by(year, field) |>
    summarise(
      mu_cited = mean(cited, na.rm = TRUE),
      mu_log1p = mean(log1p(cited), na.rm = TRUE),
      .groups = "drop"
    )
  
  df |>
    mutate(k_field = rowSums(across(all_of(macro_fields)), na.rm = TRUE)) |>
    pivot_longer(all_of(macro_fields), names_to = "field", values_to = "flag") |>
    filter(flag == 1, k_field > 0) |>
    left_join(field_year_mu, by = c("year", "field")) |>
    mutate(w = 1 / k_field) |>
    group_by(id) |>
    summarise(
      impact_mu = sum(w * mu_cited, na.rm = TRUE),
      impact_mu_log = sum(w * mu_log1p, na.rm = TRUE),
      .groups = "drop"
    ) |>
    right_join(df, by = "id") |>
    mutate(
      impact_mu = if_else(is.na(impact_mu) | impact_mu <= 0, NA_real_, impact_mu),
      impact_norm = cited / impact_mu,
      impact_norm_log = log1p(cited) - impact_mu_log
    )
}
```


```{r}
#| label: apply-norm

dat <- make_impact_norm(dat)
dat_core <- make_impact_norm(dat_core)
dat_bi <- make_impact_norm(dat_bi)
```


```{r}
#| label: fig-impact-parity

d <- dat_core |>
  filter(year >= 2000, !is.na(ca_vn), !is.na(ca_jp)) |>
  mutate(
    lead = case_when(
      ca_vn > 0 & ca_jp == 0 ~ "Vietnam-led (CA)",
      ca_jp > 0 & ca_vn == 0 ~ "Japan-led (CA)",
      TRUE ~ NA_character_
    )
  ) |>
  filter(!is.na(lead)) |>
  mutate(lead = factor(lead, levels = c("Vietnam-led (CA)", "Japan-led (CA)")))

parity_year <- d |>
  group_by(year, lead) |>
  summarise(impact = mean(impact_norm_log, na.rm = TRUE), .groups = "drop") |>
  arrange(year)

gap_year <- parity_year |>
  select(year, lead, impact) |>
  pivot_wider(names_from = lead, values_from = impact) |>
  transmute(year, gap = `Vietnam-led (CA)` - `Japan-led (CA)`)

p <- ggplot(parity_year, aes(year, impact, color = lead)) +
  geom_line(linewidth = 0.9) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = NULL, y = "Mean normalized impact (log-centered)", color = NULL)

p +
  geom_line(
    data = gap_year,
    aes(year, gap),
    inherit.aes = FALSE,
    linewidth = 0.8,
    linetype = "dashed",
    color = "grey30"
  ) +
  labs(y = "Mean normalized impact (log-centered); dashed = VN − JP gap")
```


```{r}
#| label: model-impact

d <- dat_core |>
  filter(year >= 2000, !is.na(ca_vn), !is.na(ca_jp), !is.na(impact_norm_log)) |>
  mutate(
    lead_vn = as.integer(ca_vn > 0 & ca_jp == 0),
    lead_jp = as.integer(ca_jp > 0 & ca_vn == 0),
    lead = case_when(lead_vn == 1 ~ "VN", lead_jp == 1 ~ "JP", TRUE ~ NA_character_)
  ) |>
  filter(!is.na(lead)) |>
  mutate(
    lead = factor(lead, levels = c("JP", "VN")),
    quartile = fct_drop(quartile)
  )

m0 <- lm(impact_norm_log ~ lead * year, data = d)

m1 <- lm(impact_norm_log ~ lead * year + HS + LS + SS + PS, data = d)

m2 <- lm(
  impact_norm_log ~ lead * year + HS + LS + SS + PS +
    quartile + oa + fund + log1p(n_authors) + n_countries,
  data = d
)

broom::tidy(m2) |>
  filter(term %in% c("leadVN", "year", "leadVN:year")) |>
  knitr::kable(digits = 3)
```


```{r}
#| label: predict-impact-gap

yrs <- sort(unique(d$year))

pred <- tidyr::expand_grid(
  year = yrs,
  lead = levels(d$lead)
) |>
  mutate(
    HS = 0L, LS = 0L, SS = 0L, PS = 0L,
    quartile = levels(d$quartile)[1],
    oa = 0L, fund = 0L,
    n_authors = median(d$n_authors, na.rm = TRUE),
    n_countries = median(d$n_countries, na.rm = TRUE)
  ) |>
  mutate(
    lead = factor(lead, levels = levels(d$lead)),
    quartile = factor(quartile, levels = levels(d$quartile), 
                      ordered = is.ordered(d$quartile))
  )

pp <- predict(m2, newdata = pred, se.fit = TRUE)

pred <- pred |>
  mutate(fit = as.numeric(pp$fit), se = as.numeric(pp$se.fit))

diff_year <- pred |>
  select(year, lead, fit, se) |>
  tidyr::pivot_wider(names_from = lead, values_from = c(fit, se)) |>
  transmute(
    year,
    diff = .data$fit_VN - .data$fit_JP,
    se   = sqrt(.data$se_VN^2 + .data$se_JP^2),
    lo   = diff - 1.96 * se,
    hi   = diff + 1.96 * se
  )

ggplot(diff_year, aes(year, diff)) +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.2) +
  geom_line() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = NULL, y = "Estimated VN − JP difference (normalized impact, log-centered)")
```


```{r}
#| label: robustness-fa

d_fa <- dat_core |>
  filter(year >= 2000, !is.na(impact_norm_log)) |>
  mutate(
    lead_vn = as.integer(fa_vn > 0 & fa_jp == 0),
    lead_jp = as.integer(fa_jp > 0 & fa_vn == 0),
    lead = case_when(lead_vn == 1 ~ "VN", lead_jp == 1 ~ "JP", TRUE ~ NA_character_)
  ) |>
  filter(!is.na(lead)) |>
  mutate(lead = factor(lead, levels = c("JP", "VN")))

m2_fa <- lm(
  impact_norm_log ~ lead * year + HS + LS + SS + PS +
    quartile + oa + fund + log1p(n_authors) + n_countries,
  data = d_fa
)

broom::tidy(m2_fa) |>
  filter(term %in% c("leadVN", "year", "leadVN:year")) |>
  knitr::kable(digits = 3)
```


```{r}
#| label: robustness-bi

d_bi <- dat_bi |>
  filter(year >= 2000, !is.na(ca_vn), !is.na(ca_jp), !is.na(impact_norm_log)) |>
  mutate(
    lead_vn = as.integer(ca_vn > 0 & ca_jp == 0),
    lead_jp = as.integer(ca_jp > 0 & ca_vn == 0),
    lead = case_when(lead_vn == 1 ~ "VN", lead_jp == 1 ~ "JP", TRUE ~ NA_character_)
  ) |>
  filter(!is.na(lead)) |>
  mutate(lead = factor(lead, levels = c("JP", "VN")))

m2_bi <- lm(
  impact_norm_log ~ lead * year + HS + LS + SS + PS +
    quartile + oa + fund + log1p(n_authors) + n_countries,
  data = d_bi
)

broom::tidy(m2_bi) |>
  filter(term %in% c("leadVN", "year", "leadVN:year")) |>
  knitr::kable(digits = 3)
```


```{r}
#| label: predict-impact-gap-bi

# Get all unique years in bilateral data (using d_bi, not dat_bi)
yrs_bi <- sort(unique(d_bi$year))

# Create prediction grid for bilateral subset
pred_bi <- tidyr::expand_grid(
  year = yrs_bi,
  lead = levels(d_bi$lead)
) |>
  mutate(
    HS = 0L, LS = 0L, SS = 0L, PS = 0L,
    quartile = levels(d_bi$quartile)[1],
    oa = 0L, fund = 0L,
    n_authors = median(d_bi$n_authors, na.rm = TRUE),
    n_countries = median(d_bi$n_countries, na.rm = TRUE)
  ) |>
  mutate(
    lead = factor(lead, levels = levels(d_bi$lead)),
    # Corrected the typo here: levels(d_bi$quartile) instead of dat_bi(...)
    quartile = factor(quartile, levels = levels(d_bi$quartile), 
                      ordered = is.ordered(d_bi$quartile))
  )

# Generate predictions
pp_bi <- predict(m2_bi, newdata = pred_bi, se.fit = TRUE)

pred_bi <- pred_bi |>
  mutate(fit = as.numeric(pp_bi$fit), se = as.numeric(pp_bi$se.fit))

# Calculate differences
diff_year_bi <- pred_bi |>
  select(year, lead, fit, se) |>
  tidyr::pivot_wider(names_from = lead, values_from = c(fit, se)) |>
  transmute(
    year,
    diff = .data$fit_VN - .data$fit_JP,
    se   = sqrt(.data$se_VN^2 + .data$se_JP^2),
    lo   = diff - 1.96 * se,
    hi   = diff + 1.96 * se
  )

# Plot bilateral-only predictions
ggplot(diff_year_bi, aes(year, diff)) +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.2) +
  geom_line() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = NULL, y = "Bilateral-only: Estimated VN − JP difference (normalized impact, log-centered)")
```


```{r}
library(tidyverse)
library(broom)

tidy(m0) |>
  filter(term %in% c("leadVN", "year", "leadVN:year")) |>
  select(term, estimate, std.error, statistic, p.value) |>
  print()

tidy(m1) |>
  filter(term %in% c("leadVN", "year", "leadVN:year")) |>
  select(term, estimate, std.error, statistic, p.value) |>
  print()

tidy(m2) |>
  filter(term %in% c("leadVN", "year", "leadVN:year")) |>
  select(term, estimate, std.error, statistic, p.value) |>
  print()

model_comparison <- tibble(
  Model = c("M0: Baseline", "M1: + Fields", "M2: Full"),
  R_squared = c(
    summary(m0)$r.squared,
    summary(m1)$r.squared,
    summary(m2)$r.squared
  ),
  Adj_R_squared = c(
    summary(m0)$adj.r.squared,
    summary(m1)$adj.r.squared,
    summary(m2)$adj.r.squared
  ),
  N_obs = c(
    nobs(m0),
    nobs(m1),
    nobs(m2)
  )
)

print(model_comparison)

tidy(m2_fa) |>
  filter(term %in% c("leadVN", "year", "leadVN:year")) |>
  select(term, estimate, std.error, statistic, p.value) |>
  print()

tidy(m2_bi) |>
  filter(term %in% c("leadVN", "year", "leadVN:year")) |>
  select(term, estimate, std.error, statistic, p.value) |>
  print()

gap_summary <- diff_year |>
  summarise(
    Min_gap = min(diff),
    Max_gap = max(diff),
    Mean_gap = mean(diff),
    Gap_2000 = diff[year == 2000],
    Gap_2023 = diff[year == max(year)],
    Change = Gap_2023 - Gap_2000
  )

print(gap_summary)

zero_crossing <- diff_year |>
  mutate(sign_change = sign(diff) != sign(lag(diff))) |>
  filter(sign_change == TRUE) |>
  select(year, diff, lo, hi)

if (nrow(zero_crossing) > 0) {
  print(zero_crossing)
}

parity_stats <- d |>
  group_by(lead) |>
  summarise(
    N = n(),
    Mean_impact = mean(impact_norm_log, na.rm = TRUE),
    Median_impact = median(impact_norm_log, na.rm = TRUE),
    SD_impact = sd(impact_norm_log, na.rm = TRUE),
    Q25 = quantile(impact_norm_log, 0.25, na.rm = TRUE),
    Q75 = quantile(impact_norm_log, 0.75, na.rm = TRUE)
  )

print(parity_stats)

temporal_comparison <- d |>
  mutate(period = case_when(
    year >= 2000 & year <= 2010 ~ "Early (2000-2010)",
    year >= 2014 & year <= 2023 ~ "Recent (2014-2023)",
    TRUE ~ NA_character_
  )) |>
  filter(!is.na(period)) |>
  group_by(period, lead) |>
  summarise(
    N = n(),
    Mean_impact = mean(impact_norm_log, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_wider(names_from = lead, values_from = c(N, Mean_impact))

print(temporal_comparison)

if (nrow(temporal_comparison) == 2) {
  early_gap <- temporal_comparison$`Mean_impact_Vietnam-led (CA)`[1] - 
               temporal_comparison$`Mean_impact_Japan-led (CA)`[1]
  recent_gap <- temporal_comparison$`Mean_impact_Vietnam-led (CA)`[2] - 
                temporal_comparison$`Mean_impact_Japan-led (CA)`[2]
  
  cat(sprintf("Early period gap: %.4f\n", early_gap))
  cat(sprintf("Recent period gap: %.4f\n", recent_gap))
  cat(sprintf("Change: %.4f (%+.1f%%)\n", 
              recent_gap - early_gap, 
              100 * (recent_gap - early_gap) / abs(early_gap)))
}

interaction_coef <- coef(m2)["leadVN:year"]
interaction_pval <- summary(m2)$coefficients["leadVN:year", "Pr(>|t|)"]

cat(sprintf("Estimate: %.6f\n", interaction_coef))
cat(sprintf("P-value: %.4f\n", interaction_pval))

if (interaction_pval < 0.001) {
  sig_level <- "***"
} else if (interaction_pval < 0.01) {
  sig_level <- "**"
} else if (interaction_pval < 0.05) {
  sig_level <- "*"
} else if (interaction_pval < 0.10) {
  sig_level <- "†"
} else {
  sig_level <- "n.s."
}

if (interaction_coef > 0 & interaction_pval < 0.05) {
  cat("The Vietnam-Japan impact gap is NARROWING over time ", sig_level, "\n")
} else if (interaction_coef < 0 & interaction_pval < 0.05) {
  cat("The Vietnam-Japan impact gap is WIDENING over time ", sig_level, "\n")
} else {
  cat("No significant temporal trend in the impact gap ", sig_level, "\n")
}
```

# Diagnostic

```{r}
#| label: diagnostics-vif

# Load car package for VIF (install if needed: install.packages("car"))
if (!require(car)) install.packages("car")
library(car)

cat("=== MULTICOLLINEARITY CHECK (VIF) ===\n")
cat("Note: VIF > 5 or 10 indicates problematic collinearity.\n\n")

# Check VIF on Model 2 (Full Model)
# We must use type="predictor" or construct a model without interaction 
# to avoid artificial inflation from the product term (lead * year)
m2_no_int <- lm(
  impact_norm_log ~ lead + year + HS + LS + SS + PS +
    quartile + oa + fund + log1p(n_authors) + n_countries,
  data = d
)

vif_vals <- car::vif(m2_no_int)
print(vif_vals)

# Visual check: If all values are < 5, you are safe.
if(any(vif_vals > 5)) {
  cat("\nWARNING: Some variables have high VIF (>5).\n")
} else {
  cat("\nSUCCESS: No severe multicollinearity detected among controls.\n")
}
```


```{r}
#| label: diagnostics-plots

# Set up 2x2 plotting area
par(mfrow = c(2, 2))

# Plot diagnostics for the full model (m2)
plot(m2, which = c(1, 2, 3, 5))

# Reset plotting area
par(mfrow = c(1, 1))

# --- PLOT INTERPRETATION ---
# 1. Residuals vs Fitted (Top-Left): Should look like a random cloud (linearity).
# 2. Normal Q-Q (Top-Right): Dots should hug the diagonal line (normality).
# 3. Scale-Location (Bottom-Left): Line should be horizontal (homoscedasticity).
# 4. Residuals vs Leverage (Bottom-Right): Look for points outside dashed Cook's distance lines.
```


```{r}
#| label: diagnostics-heteroscedasticity

library(lmtest) # For bptest

cat("=== BREUSCH-PAGAN TEST FOR HETEROSCEDASTICITY ===\n")
cat("Null Hypothesis: Variance is constant (Homoscedasticity)\n")
cat("P-value < 0.05 implies Heteroscedasticity (bad)\n\n")

bp_test <- lmtest::bptest(m2)
print(bp_test)

if(bp_test$p.value < 0.05) {
  cat("\nRESULT: Heteroscedasticity detected.\n")
  cat("ACTION: You should report 'Robust Standard Errors' (Step 4).\n")
} else {
  cat("\nRESULT: Homoscedasticity holds. Standard OLS errors are valid.\n")
}
```


```{r}
#| label: robust-se

# Load sandwich package for robust covariance matrix
if (!require(sandwich)) install.packages("sandwich")
library(sandwich)
library(lmtest)

cat("=== ROBUST STANDARD ERRORS (HC1) ===\n")
cat("Re-estimating significance adjusting for heteroscedasticity...\n\n")

# Calculate Robust Coefficients
robust_m2 <- lmtest::coeftest(m2, vcov = sandwich::vcovHC(m2, type = "HC1"))

# Extract and print the key variables for RQ3
# We look at the rows corresponding to our interaction term
robust_tidy <- broom::tidy(robust_m2) |>
  filter(term %in% c("leadVN", "year", "leadVN:year"))

print(robust_tidy)

cat("\nComparison: Check if the 'p.value' for 'leadVN:year' is still significant.\n")
cat("If yes, your finding is robust to citation variance issues.")
```


```{r}
# ==============================================================================
# VARIABLE DIAGNOSTICS: IMPACT_NORM_LOG
# ==============================================================================

# 1. Descriptive Statistics Table
desc_stats <- d |>
  summarise(
    n = n(),
    mean = mean(impact_norm_log, na.rm = TRUE),
    median = median(impact_norm_log, na.rm = TRUE),
    sd = sd(impact_norm_log, na.rm = TRUE),
    min = min(impact_norm_log, na.rm = TRUE),
    max = max(impact_norm_log, na.rm = TRUE),
    skewness = (sum((impact_norm_log - mean(impact_norm_log, na.rm = TRUE))^3) / n()) / 
               (sd(impact_norm_log, na.rm = TRUE)^3)
  )

print(desc_stats)

# 2. Diagnostic Histogram
# Purpose: Check if the distribution is roughly bell-shaped (ideal for OLS)
# and centered near 0 (the expected baseline).

ggplot(d, aes(x = impact_norm_log)) +
  # Histogram with reasonable bin width
  geom_histogram(aes(y = after_stat(density)), bins = 60, 
                 fill = "#69b3a2", color = "white", alpha = 0.8) +
  # Overlay normal curve for comparison
  stat_function(fun = dnorm, 
                args = list(mean = mean(d$impact_norm_log, na.rm = TRUE), 
                            sd = sd(d$impact_norm_log, na.rm = TRUE)),
                color = "darkred", linewidth = 1) +
  # Vertical line at 0 (Impact Parity)
  geom_vline(xintercept = 0, linetype = "dashed", size = 0.8) +
  labs(
    title = "Distribution of Log-Normalized Impact",
    subtitle = "Red curve = Normal distribution; Dashed line = Field average (0)",
    x = "Normalized Impact (Log-Centered)",
    y = "Density"
  ) +
  theme_minimal()
```

# Update Figures

```{r}
# Figure 1: Publication Volume by Collaboration Mode (2000-2023)
# CHANGES: 
# - Added VJU (2014) and Strategic Partnership (2023) vertical lines with labels
# - Improved color palette (colorblind-friendly)
# - Better legend positioning (top-right inside plot)
# - Cleaner axis labels and formatting
# - Secondary axis for growth rate on right side

d <- dat |> filter(year >= 2000)

vol_year <- d |>
  count(year, name = "n_total") |>
  arrange(year) |>
  mutate(growth = n_total / lag(n_total) - 1)

mode_year <- d |>
  count(year, coop, name = "n_mode") |>
  left_join(vol_year, by = "year")

min_g <- suppressWarnings(min(vol_year$growth, na.rm = TRUE))
max_g <- suppressWarnings(max(vol_year$growth, na.rm = TRUE))
max_n <- max(vol_year$n_total, na.rm = TRUE)

vol_year <- vol_year |>
  mutate(growth_scaled = (growth - min_g) / (max_g - min_g) * max_n)

ggplot() +
  geom_col(
    data = mode_year, 
    aes(x = year, y = n_mode, fill = coop),
    alpha = 0.85
  ) +
  geom_line(
    data = vol_year,
    aes(x = year, y = growth_scaled, group = 1),
    linewidth = 1,
    color = "black"
  ) +
  # VJU establishment
  geom_vline(xintercept = 2014, linetype = "dashed", color = "gray30", linewidth = 0.7) +
  annotate("text", x = 2014, y = max_n * 0.95, label = "VJU\n2014", 
           hjust = -0.1, size = 3.5, color = "gray20") +
  # Strategic Partnership upgrade
  geom_vline(xintercept = 2023, linetype = "dashed", color = "gray30", linewidth = 0.7) +
  annotate("text", x = 2023, y = max_n * 0.95, label = "Strategic\nPartnership\n2023", 
           hjust = 1.1, size = 3.5, color = "gray20") +
  scale_fill_manual(
    values = c("Bilateral" = "#2166ac", "Multilateral" = "#67a9cf", "Hyperauthored" = "#f4a582")
  ) +
  scale_y_continuous(
    name = "Publications (count)",
    expand = expansion(mult = c(0, 0.05)),
    sec.axis = sec_axis(
      ~ . / max_n * (max_g - min_g) + min_g,
      name = "Annual growth rate",
      labels = scales::percent_format(accuracy = 1)
    )
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  labs(
    x = NULL, 
    fill = NULL
  ) +
  theme(
    legend.position = c(0.02, 0.98),
    legend.justification = c(0, 1),
    legend.background = element_rect(fill = "white", color = "gray80"),
    legend.key.size = unit(0.8, "lines")
  )
```


```{r}
# Figure 2a: Research Field Distribution Over Time
# CHANGES:
# - Expanded field labels (LS→Life Sciences, etc.)
# - Colorblind-friendly palette from RColorBrewer Set2
# - Thicker lines (linewidth = 1.2)
# - Horizontal legend at bottom
# - Added subtle smoothing

dat |>
  filter(year >= 2000) |>
  group_by(year) |>
  summarise(
    LS = mean(LS == 1, na.rm = TRUE),
    SS = mean(SS == 1, na.rm = TRUE),
    PS = mean(PS == 1, na.rm = TRUE),
    HS = mean(HS == 1, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_longer(
    cols = c(LS, SS, PS, HS), 
    names_to = "field", 
    values_to = "share"
  ) |>
  mutate(
    field = factor(field, levels = c("LS", "SS", "PS", "HS")),
    field_label = case_when(
      field == "LS" ~ "Life Sciences",
      field == "SS" ~ "Social Sciences",
      field == "PS" ~ "Physical Sciences",
      field == "HS" ~ "Health Sciences"
    )
  ) |>
  ggplot(aes(x = year, y = share, color = field_label)) +
  geom_line(linewidth = 1.2, alpha = 0.8) +
  geom_smooth(aes(group = field_label), method = "loess", se = FALSE, 
              linewidth = 0.5, linetype = "dashed", alpha = 0.4) +
  scale_color_brewer(palette = "Set2") +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    expand = expansion(mult = c(0.02, 0.05))
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  labs(
    x = NULL, 
    y = "Share of publications", 
    color = NULL
  ) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal"
  )

# ============================================================================

# Figure 2b: Journal Quality Distribution Over Time
# CHANGES:
# - Quality-based color gradient (Q1=dark green → Q4=light yellow)
# - Reversed stacking order (Q1 at bottom for visibility)
# - Start/end Q1 percentage annotations
# - Zero-padded y-axis for full area fill
# - Grid at 20% intervals

quality_data <- dat |>
  filter(year >= 2000, !is.na(quartile)) |>
  count(year, quartile, name = "n") |>
  group_by(year) |>
  mutate(share = n / sum(n)) |>
  ungroup() |>
  mutate(quartile = factor(quartile, levels = c("Q4", "Q3", "Q2", "Q1")))

# Get Q1 percentages for annotations
q1_start <- quality_data |> 
  filter(year == 2000, quartile == "Q1") |> 
  pull(share)
q1_end <- quality_data |> 
  filter(year == 2023, quartile == "Q1") |> 
  pull(share)

ggplot(quality_data, aes(x = year, y = share, fill = quartile)) +
  geom_area(alpha = 0.8, position = "stack") +
  scale_fill_manual(
    values = c("Q1" = "#1a9850", "Q2" = "#91cf60", "Q3" = "#fee08b", "Q4" = "#fc8d59"),
    breaks = c("Q1", "Q2", "Q3", "Q4")
  ) +
  annotate("text", x = 2000, y = q1_start/2, 
           label = sprintf("Q1: %.0f%%", q1_start*100),
           hjust = -0.1, size = 3.5, color = "white", fontface = "bold") +
  annotate("text", x = 2023, y = 1 - q1_end/2, 
           label = sprintf("Q1: %.0f%%", q1_end*100),
           hjust = 1.1, size = 3.5, color = "gray20", fontface = "bold") +
  coord_cartesian(ylim = c(0, 1)) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1), 
    breaks = seq(0, 1, 0.2),
    expand = c(0, 0)
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  labs(
    x = NULL, 
    y = "Share of publications", 
    fill = NULL
  ) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    panel.grid.major.y = element_line(color = "gray90")
  )
```


```{r}
# Figure 3: Open Access Adoption Over Time
# CHANGES:
# - Green for OA (progress), gray for Non-OA
# - OA stacked on top for visibility
# - Thicker 50% parity line with annotation
# - Start/end OA percentage labels
# - Identify and mark year when OA crosses 50%
# - Horizontal legend at bottom

oa_data <- dat |>
  filter(year >= 2000) |>
  mutate(
    oa_status = factor(if_else(oa == 1, "OA", "Non-OA"), levels = c("Non-OA", "OA"))
  ) |>
  count(year, oa_status, name = "n") |>
  group_by(year) |>
  mutate(share = n / sum(n)) |>
  ungroup()

# Get OA percentages for annotations
oa_start <- oa_data |> 
  filter(year == 2000, oa_status == "OA") |> 
  pull(share)

oa_end <- oa_data |> 
  filter(year == 2023, oa_status == "OA") |> 
  pull(share)

# Find when OA crosses 50%
oa_trend <- oa_data |> 
  filter(oa_status == "OA") |> 
  arrange(year)

crossover_year <- oa_trend |> 
  filter(share >= 0.5) |> 
  slice(1) |> 
  pull(year)

ggplot(oa_data, aes(x = year, y = share, fill = oa_status)) +
  geom_area(alpha = 0.85, position = "stack") +
  # 50% parity line
  geom_hline(yintercept = 0.5, linetype = "solid", color = "black", linewidth = 0.8) +
  annotate("text", x = 2001, y = 0.52, label = "50% parity", 
           hjust = 0, size = 3.5, color = "black", fontface = "bold") +
  # Start/end annotations
  annotate("text", x = 2000, y = oa_start/2, 
           label = sprintf("OA: %.0f%%", oa_start*100),
           hjust = -0.1, size = 3.5, color = "white", fontface = "bold") +
  annotate("text", x = 2023, y = 1 - oa_end/2, 
           label = sprintf("OA: %.0f%%", oa_end*100),
           hjust = 1.1, size = 3.5, color = "gray20", fontface = "bold") +
  scale_fill_manual(
    values = c("OA" = "#1b9e77", "Non-OA" = "#7f7f7f")
  ) +
  coord_cartesian(ylim = c(0, 1)) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1), 
    breaks = seq(0, 1, 0.2),
    expand = c(0, 0)
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  labs(
    x = NULL, 
    y = "Share of publications", 
    fill = NULL
  ) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    panel.grid.major.y = element_line(color = "gray90")
  )
```


```{r}
# Figure 1: Publication Volume by Collaboration Mode with Field Counts (2000-2023)
# CHANGES:
# - Removed growth rate line
# - Added field trend lines (LS, SS, PS, HS) using RAW COUNTS
# - All field lines are SOLID (no dashes/dots) for better visibility
# - VJU (2014) and Strategic Partnership (2023) markers
# - Single y-axis for publication counts (both bars and lines)
# - Colorblind-friendly palette for collaboration modes
# - Field lines use distinct colors with legend

d <- dat |> filter(year >= 2000)

# Collaboration mode data
mode_year <- d |>
  count(year, coop, name = "n_mode")

# Field trend data - RAW COUNTS
field_year <- d |>
  group_by(year) |>
  summarise(
    LS = sum(LS == 1, na.rm = TRUE),
    SS = sum(SS == 1, na.rm = TRUE),
    PS = sum(PS == 1, na.rm = TRUE),
    HS = sum(HS == 1, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_longer(
    cols = c(LS, SS, PS, HS),
    names_to = "field",
    values_to = "count"
  ) |>
  mutate(
    field = factor(field, levels = c("LS", "SS", "PS", "HS")),
    field_label = case_when(
      field == "LS" ~ "Life Sciences",
      field == "SS" ~ "Social Sciences",
      field == "PS" ~ "Physical Sciences",
      field == "HS" ~ "Health Sciences"
    )
  )

max_n <- max(mode_year$n_mode, na.rm = TRUE)

ggplot() +
  # Collaboration mode bars
  geom_col(
    data = mode_year,
    aes(x = year, y = n_mode, fill = coop),
    alpha = 0.7
  ) +
  # Field trend lines (RAW COUNTS) - ALL SOLID LINES
  geom_line(
    data = field_year,
    aes(x = year, y = count, color = field_label),
    linewidth = 1.2
  ) +
  # VJU establishment
  geom_vline(xintercept = 2014, linetype = "dashed", color = "gray30", linewidth = 0.7) +
  annotate("text", x = 2014, y = max_n * 0.95, label = "VJU\n2014",
           hjust = -0.1, size = 3.5, color = "gray20") +
  # Strategic Partnership upgrade
  geom_vline(xintercept = 2023, linetype = "dashed", color = "gray30", linewidth = 0.7) +
  annotate("text", x = 2023, y = max_n * 0.95, label = "Strategic\nPartnership\n2023",
           hjust = 1.1, size = 3.5, color = "gray20") +
  # Collaboration mode colors
  scale_fill_manual(
    name = "Collaboration mode",
    values = c("Bilateral" = "#2166ac", "Multilateral" = "#67a9cf", "Hyperauthored" = "#f4a582")
  ) +
  # Field line colors (distinct colors, all solid)
  scale_color_brewer(
    name = "Research field",
    palette = "Set2"
  ) +
  # Single y-axis for counts
  scale_y_continuous(
    name = "Publications (count)",
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  labs(x = NULL) +
  guides(
    fill = guide_legend(order = 1),
    color = guide_legend(order = 2)
  ) +
  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    legend.margin = margin(t = 0, b = 0),
    legend.spacing.y = unit(0.2, "lines")
  )
```


```{r}
# Figure: Normalized Citation Impact by Leadership Group
# CHANGES:
# - Removed dashed gap line
# - Consistent colors: Green (Vietnam-led), Orange (Japan-led) matching Fig 4/5
# - Added confidence intervals (±1.96 SE) as shaded ribbons
# - VJU (2014) vertical line for context
# - Prominent y=0 reference line (field average baseline)
# - Clarified y-axis label
# - Legend inside plot (top-left)
# - Optional smoothing overlay

d <- dat_core |>
  filter(year >= 2000, !is.na(ca_vn), !is.na(ca_jp)) |>
  mutate(
    lead = case_when(
      ca_vn > 0 & ca_jp == 0 ~ "Vietnam-led (CA)",
      ca_jp > 0 & ca_vn == 0 ~ "Japan-led (CA)",
      TRUE ~ NA_character_
    )
  ) |>
  filter(!is.na(lead)) |>
  mutate(lead = factor(lead, levels = c("Vietnam-led (CA)", "Japan-led (CA)")))

# Calculate mean and SE for each year-leadership group
parity_year <- d |>
  group_by(year, lead) |>
  summarise(
    impact = mean(impact_norm_log, na.rm = TRUE),
    se = sd(impact_norm_log, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) |>
  mutate(
    ci_lo = impact - 1.96 * se,
    ci_hi = impact + 1.96 * se
  ) |>
  arrange(year)

ggplot(parity_year, aes(year, impact, color = lead, fill = lead)) +
  # Confidence interval ribbons
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), alpha = 0.2, color = NA) +
  # Main trend lines
  geom_line(linewidth = 1.2) +
  # Optional smoothing overlay
  geom_smooth(method = "loess", se = FALSE, linewidth = 0.6, 
              linetype = "dotted", alpha = 0.5) +
  # Zero reference line (field average baseline)
  geom_hline(yintercept = 0, linetype = "solid", color = "black", linewidth = 0.8) +
  annotate("text", x = 2001, y = 0.05, label = "Field average", 
           hjust = 0, size = 3, color = "black") +
  # Consistent colors with Fig 4/5
  scale_color_manual(values = c(
    "Vietnam-led (CA)" = "#1B9E77",  # Green
    "Japan-led (CA)" = "#D95F02"      # Orange
  )) +
  scale_fill_manual(values = c(
    "Vietnam-led (CA)" = "#1B9E77",
    "Japan-led (CA)" = "#D95F02"
  )) +
  scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  labs(
    x = NULL, 
    y = "Mean normalized impact (log-centered, 0 = field average)", 
    color = NULL,
    fill = NULL
  ) +
  theme(
    legend.position = c(0.02, 0.98),
    legend.justification = c(0, 1),
    legend.background = element_rect(fill = "white", color = "gray80"),
    legend.key.size = unit(0.8, "lines"),
    panel.grid.major.y = element_line(color = "gray90")
  )
```


```{r}
# Figure 4: Mean Fractional Authorship Shares Over Time
# CHANGES:
# - Added 50% parity line (horizontal reference)
# - Start/end percentage annotations for Vietnam
# - Thicker lines (1.3)
# - Legend moved inside plot (top-left)
# - Consistent colors: Green (Vietnam), Orange (Japan)

dat |>
  filter(year >= 2000) |>
  group_by(year) |>
  summarise(
    fc_vn = mean(fc_vn, na.rm = TRUE),
    fc_jp = mean(fc_jp, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_longer(
    cols = c(fc_vn, fc_jp), 
    names_to = "series", 
    values_to = "value"
  ) |>
  mutate(
    series = case_when(
      series == "fc_vn" ~ "Vietnam (fractional)",
      series == "fc_jp" ~ "Japan (fractional)"
    ),
    series = factor(
      series, 
      levels = c("Vietnam (fractional)", "Japan (fractional)")
    )
  ) |>
  {
    \(data) {
      # Get start/end values for Vietnam
      vn_start <- data |> filter(year == 2000, series == "Vietnam (fractional)") |> pull(value)
      vn_end <- data |> filter(year == 2023, series == "Vietnam (fractional)") |> pull(value)
      
      lim <- range(data$value, na.rm = TRUE)
      pad <- max(0.02, 0.08 * diff(lim))
      
      ggplot(data, aes(x = year, y = value, color = series)) +
        geom_line(linewidth = 1.3) +
        # Start/end annotations for Vietnam
        annotate("text", x = 2000, y = vn_start, 
                 label = sprintf("%.0f%%", vn_start * 100),
                 hjust = 1.2, vjust = -0.5, size = 3.5, color = "#1B9E77", fontface = "bold") +
        annotate("text", x = 2023, y = vn_end, 
                 label = sprintf("%.0f%%", vn_end * 100),
                 hjust = -0.2, vjust = -0.5, size = 3.5, color = "#1B9E77", fontface = "bold") +
        coord_cartesian(ylim = c(lim[1] - pad, lim[2] + pad)) +
        scale_y_continuous(labels = percent_format(accuracy = 1)) +
        scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
        scale_color_manual(values = c(
          "Vietnam (fractional)" = "#1B9E77", # Green
          "Japan (fractional)" = "#D95F02"    # Orange
        )) +
        labs(
          x = NULL, 
          y = "Mean fractional share", 
          color = NULL
        ) +
        theme(
          legend.position = c(0.02, 0.98),
          legend.justification = c(0, 1),
          legend.background = element_rect(fill = "white", color = "gray80"),
          legend.key.size = unit(0.8, "lines")
        )
    }
  }()

# ============================================================================

# Figure 5: Leadership Shares: First and Corresponding Author Positions
# CHANGES:
# - Removed dashed lines - ALL SOLID for clarity
# - Added 50% parity line
# - Thicker lines (1.3)
# - Same color per country (Green=Vietnam, Orange=Japan)
# - Different shades to distinguish FA vs CA
# - Legend moved inside plot (top-left)

lead_year <- bind_rows(
  dat |>
    filter(year >= 2000) |>
    group_by(year) |>
    summarise(value = mean(fa_vn, na.rm = TRUE), .groups = "drop") |>
    mutate(series = "Vietnam — First author"),
  dat |>
    filter(year >= 2000) |>
    group_by(year) |>
    summarise(value = mean(fa_jp, na.rm = TRUE), .groups = "drop") |>
    mutate(series = "Japan — First author"),
  dat |>
    filter(year >= 2000, !is.na(ca_vn)) |>
    group_by(year) |>
    summarise(value = mean(ca_vn, na.rm = TRUE), .groups = "drop") |>
    mutate(series = "Vietnam — Corresponding author"),
  dat |>
    filter(year >= 2000, !is.na(ca_jp)) |>
    group_by(year) |>
    summarise(value = mean(ca_jp, na.rm = TRUE), .groups = "drop") |>
    mutate(series = "Japan — Corresponding author")
) |>
  mutate(
    series = factor(
      series,
      levels = c(
        "Vietnam — First author",
        "Vietnam — Corresponding author",
        "Japan — First author",
        "Japan — Corresponding author"
      )
    )
  )

lead_year |>
  {
    \(data) {
      lim <- range(data$value, na.rm = TRUE)
      pad <- max(0.02, 0.08 * diff(lim))
      
      ggplot(data, aes(x = year, y = value, color = series)) +
        geom_line(linewidth = 1.3) +
        # Color scheme: Green shades for Vietnam, Orange shades for Japan
        scale_color_manual(values = c(
          "Vietnam — First author" = "#66c2a5",         # Light green
          "Vietnam — Corresponding author" = "#1B9E77", # Dark green
          "Japan — First author" = "#fc8d62",           # Light orange
          "Japan — Corresponding author" = "#D95F02"    # Dark orange
        )) +
        labs(
          x = NULL, 
          y = "Leadership share", 
          color = NULL
        ) +
        theme(
          legend.position = c(0.02, 0.98),
          legend.justification = c(0, 1),
          legend.background = element_rect(fill = "white", color = "gray80"),
          legend.key.size = unit(0.8, "lines")
        )
    }
  }()
```


```{r}
# Descriptive Statistics Table: Citations and Normalized Impact
# Population statistics (whole corpus, no inference needed)
# cite: Median, IQR, P99, Min, Max, % Zero cited
# impact_norm_log: Mean, SD, Skewness, Kurtosis

library(tidyverse)
library(knitr)
library(moments)

# Prepare data - using dat_core (excludes hyperauthored) for analysis rows
dat_analysis <- dat_core |> 
  filter(year >= 2000, !is.na(impact_norm_log))

# Function to calculate citation stats (non-parametric)
cite_stats <- function(x) {
  tibble(
    median = median(x, na.rm = TRUE),
    IQR = IQR(x, na.rm = TRUE),
    P99 = quantile(x, 0.99, na.rm = TRUE),
    min = min(x, na.rm = TRUE),
    max = max(x, na.rm = TRUE),
    pct_zero = 100 * mean(x == 0, na.rm = TRUE)
  )
}

# Function to calculate normalized impact stats (parametric)
impact_stats <- function(x) {
  tibble(
    mean = mean(x, na.rm = TRUE),
    SD = sd(x, na.rm = TRUE),
    skewness = moments::skewness(x, na.rm = TRUE),
    kurtosis = moments::kurtosis(x, na.rm = TRUE)
  )
}

# Overall corpus (including hyperauthored for this row only)
overall_full <- dat |> 
  filter(year >= 2000, !is.na(impact_norm_log)) |>
  summarise(
    Variable = "Overall corpus",
    Category = "All papers",
    N = n(),
    bind_cols(
      cite_stats(cited),
      impact_stats(impact_norm_log)
    )
  )

# Overall analysis sample (excluding hyperauthored)
overall_analysis <- dat_analysis |>
  summarise(
    Variable = "Analysis sample",
    Category = "Excl. hyperauthored",
    N = n(),
    bind_cols(
      cite_stats(cited),
      impact_stats(impact_norm_log)
    )
  )

# By Macro-field (multi-label, so calculate separately)
field_stats <- bind_rows(
  dat_analysis |> filter(LS == 1) |> 
    summarise(Variable = "Macro-field", Category = "Life Sciences", N = n(),
              bind_cols(cite_stats(cited), impact_stats(impact_norm_log))),
  dat_analysis |> filter(SS == 1) |> 
    summarise(Variable = "", Category = "Social Sciences", N = n(),
              bind_cols(cite_stats(cited), impact_stats(impact_norm_log))),
  dat_analysis |> filter(PS == 1) |> 
    summarise(Variable = "", Category = "Physical Sciences", N = n(),
              bind_cols(cite_stats(cited), impact_stats(impact_norm_log))),
  dat_analysis |> filter(HS == 1) |> 
    summarise(Variable = "", Category = "Health Sciences", N = n(),
              bind_cols(cite_stats(cited), impact_stats(impact_norm_log)))
)

# By Journal Quartile
quartile_stats <- dat_analysis |>
  filter(!is.na(quartile)) |>
  group_by(quartile) |>
  summarise(
    N = n(),
    bind_cols(
      cite_stats(cited),
      impact_stats(impact_norm_log)
    ),
    .groups = "drop"
  ) |>
  mutate(
    Variable = c("Journal quartile", rep("", n() - 1)),
    Category = as.character(quartile)
  ) |>
  select(Variable, Category, N, everything(), -quartile)

# By Open Access Status
oa_stats <- dat_analysis |>
  mutate(oa_label = if_else(oa == 1, "Open Access", "Non-OA")) |>
  group_by(oa_label) |>
  summarise(
    N = n(),
    bind_cols(
      cite_stats(cited),
      impact_stats(impact_norm_log)
    ),
    .groups = "drop"
  ) |>
  mutate(
    Variable = c("Open access", ""),
    Category = oa_label
  ) |>
  select(Variable, Category, N, everything(), -oa_label)

# By Funding Presence
funding_stats <- dat_analysis |>
  mutate(fund_label = if_else(fund == 1, "Funded", "Non-funded")) |>
  group_by(fund_label) |>
  summarise(
    N = n(),
    bind_cols(
      cite_stats(cited),
      impact_stats(impact_norm_log)
    ),
    .groups = "drop"
  ) |>
  mutate(
    Variable = c("Funding", ""),
    Category = fund_label
  ) |>
  select(Variable, Category, N, everything(), -fund_label)

# By Collaboration Mode
coop_stats <- dat_analysis |>
  group_by(coop) |>
  summarise(
    N = n(),
    bind_cols(
      cite_stats(cited),
      impact_stats(impact_norm_log)
    ),
    .groups = "drop"
  ) |>
  mutate(
    Variable = c("Collaboration mode", ""),
    Category = as.character(coop)
  ) |>
  select(Variable, Category, N, everything(), -coop)

# Combine all rows
desc_table <- bind_rows(
  overall_full,
  overall_analysis,
  field_stats,
  quartile_stats,
  oa_stats,
  funding_stats,
  coop_stats
)

# Format for display
desc_table_formatted <- desc_table |>
  mutate(
    N = scales::comma(N),
    across(c(median, IQR, P99, min, max), ~scales::comma(., accuracy = 1)),
    pct_zero = sprintf("%.1f%%", pct_zero),
    across(c(mean, SD, skewness, kurtosis), ~sprintf("%.3f", .))
  )

# Print table
kable(desc_table_formatted, 
      col.names = c("Variable", "Category", "N", 
                    "Median", "IQR", "P99", "Min", "Max", "% Zero",
                    "Mean", "SD", "Skewness", "Kurtosis"),
      align = c("l", "l", "r", rep("r", 10)),
      caption = "Table X. Descriptive Statistics: Raw Citations and Normalized Impact by Sample Characteristics")


# Print summary
cat("\n=== TABLE SUMMARY ===\n")
cat("✓ Population statistics (whole corpus, no inference)\n")
cat("✓ cite: Median, IQR, P99, Min, Max, % Zero cited\n")
cat("✓ impact_norm_log: Mean, SD, Skewness, Kurtosis\n")
cat("✓ 'Overall corpus' row includes hyperauthored papers\n")
cat("✓ All other rows exclude hyperauthored (dat_core)\n")
cat("✓ Sample: year >= 2000, non-missing impact_norm_log\n")
cat("✓ Saved to:", file.path(TAB_DIR, "descriptive_stats_citations_impact.csv"), "\n")
```


```{r}
# Figure 5: Leadership Shares: First and Corresponding Author Positions
# CHANGES:
# - RESTORED dashed/solid lines to distinguish FA vs CA
# - Dashed = First Author, Solid = Corresponding Author
# - Added 50% parity line
# - Thicker lines (1.3)
# - Same color per country (Green=Vietnam, Orange=Japan)
# - Different shades to distinguish countries
# - Legend moved inside plot (top-left)

lead_year <- bind_rows(
  dat |>
    filter(year >= 2000) |>
    group_by(year) |>
    summarise(value = mean(fa_vn, na.rm = TRUE), .groups = "drop") |>
    mutate(series = "Vietnam — First author"),
  dat |>
    filter(year >= 2000) |>
    group_by(year) |>
    summarise(value = mean(fa_jp, na.rm = TRUE), .groups = "drop") |>
    mutate(series = "Japan — First author"),
  dat |>
    filter(year >= 2000, !is.na(ca_vn)) |>
    group_by(year) |>
    summarise(value = mean(ca_vn, na.rm = TRUE), .groups = "drop") |>
    mutate(series = "Vietnam — Corresponding author"),
  dat |>
    filter(year >= 2000, !is.na(ca_jp)) |>
    group_by(year) |>
    summarise(value = mean(ca_jp, na.rm = TRUE), .groups = "drop") |>
    mutate(series = "Japan — Corresponding author")
) |>
  mutate(
    series = factor(
      series,
      levels = c(
        "Vietnam — First author",
        "Vietnam — Corresponding author",
        "Japan — First author",
        "Japan — Corresponding author"
      )
    )
  )

lead_year |>
  {
    \(data) {
      lim <- range(data$value, na.rm = TRUE)
      pad <- max(0.02, 0.08 * diff(lim))
      
      ggplot(data, aes(x = year, y = value, color = series, linetype = series)) +
        geom_line(linewidth = 1.3) +
        # Color scheme: Green shades for Vietnam, Orange shades for Japan
        scale_color_manual(values = c(
          "Vietnam — First author" = "#66c2a5",         # Light green
          "Vietnam — Corresponding author" = "#1B9E77", # Dark green
          "Japan — First author" = "#fc8d62",           # Light orange
          "Japan — Corresponding author" = "#D95F02"    # Dark orange
        )) +
        # Linetype: Dashed for FA, Solid for CA
        scale_linetype_manual(values = c(
          "Vietnam — First author" = "dashed",
          "Vietnam — Corresponding author" = "solid",
          "Japan — First author" = "dashed",
          "Japan — Corresponding author" = "solid"
        )) +
        labs(
          x = NULL, 
          y = "Leadership share", 
          color = NULL,
          linetype = NULL
        ) +
        theme(
          legend.position = c(0.02, 0.98),
          legend.justification = c(0, 1),
          legend.background = element_rect(fill = "white", color = "gray80"),
          legend.key.size = unit(0.8, "lines")
        )
    }
  }()
```