---
title: "Changing in Equity and Leadership in Japan–Vietnam Scientific Collaboration"
subtitle: "A Bibliometric Assessment — Analysis"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    df-print: paged
    fig-width: 9
    fig-height: 5
execute:
  echo: true
  warning: false
  message: false
  cache: false
  freeze: auto
---

# Setup

This section initializes the analysis environment:

- Loads required packages for data handling and graphics.

- Sets global options for reproducible, consistent outputs.

- Defines project paths and creates output directories if missing.

- Establishes a unified ggplot theme (fonts, sizes, spacing) to keep figures visually consistent.


```{r}
#| label: setup
#| include: false

library(tidyverse)
library(readr)
library(scales)
library(lubridate)

set.seed(20260105)

options(
  dplyr.summarise.inform = FALSE,
  readr.show_col_types = FALSE,
  stringsAsFactors = FALSE
)

DATA_DIR <- "data"
OUT_DIR  <- "outputs"
FIG_DIR  <- file.path(OUT_DIR, "figures")
TAB_DIR  <- file.path(OUT_DIR, "tables")

dir.create(OUT_DIR, showWarnings = FALSE, recursive = TRUE)
dir.create(FIG_DIR, showWarnings = FALSE, recursive = TRUE)
dir.create(TAB_DIR, showWarnings = FALSE, recursive = TRUE)

theme_set(
  theme_minimal(base_size = 12, base_family = "sans") +
    theme(
      plot.title.position   = "plot",
      plot.caption.position = "plot",
      legend.position       = "bottom",
      legend.title          = element_text(size = 10),
      legend.text           = element_text(size = 10),
      axis.title            = element_text(size = 11),
      axis.text             = element_text(size = 10),
      strip.text            = element_text(size = 11),
      panel.grid.minor      = element_blank()
    )
)

ggplot2::update_geom_defaults("line",   list(linewidth = 0.9))
ggplot2::update_geom_defaults("smooth", list(linewidth = 0.9))
ggplot2::update_geom_defaults("col",    list(linewidth = 0))
ggplot2::update_geom_defaults("area",   list(linewidth = 0))
```

# Data and samples

```{r}
#| label: load-data

DATA_FILE <- file.path(DATA_DIR, "updated_data.csv")

dat <- read_csv(DATA_FILE) |>
  mutate(
    year = as.integer(year),
    coop = factor(coop, levels = c("Bilateral", "Multilateral", "Hyperauthored")),
    quartile = factor(quartile, levels = c("Q1", "Q2", "Q3", "Q4"), ordered = TRUE),
    oa = as.integer(oa),
    fund = as.integer(fund),
    across(c(LS, SS, PS, HS), as.integer)
  ) |>
  filter(!is.na(coop))

dat_core <- dat |> filter(coop != "Hyperauthored")

dat_bi <- dat |> filter(coop == "Bilateral")
```


```{r}
#| label: tbl-profile

fmt_iqr <- function(x) {
  qs <- quantile(x, probs = c(0.25, 0.50, 0.75), na.rm = TRUE, type = 2)
  paste0(scales::comma(qs[[2]]), " (", scales::comma(qs[[1]]), "–", scales::comma(qs[[3]]), ")")
}

fmt_n_pct <- function(n, d) {
  if (is.na(d) || d == 0) return(paste0(scales::comma(n), " (NA)"))
  paste0(scales::comma(n), " (", sprintf("%.1f", 100 * n / d), "%)")
}

mode_stats <- bind_rows(
  dat |> mutate(mode = as.character(coop)),
  dat |> mutate(mode = "Total")
  ) |>
  group_by(mode) |>
  summarise(
    n = n(),
    n_auth_iqr = fmt_iqr(n_authors),
    n_ctry_iqr = fmt_iqr(n_countries),
    oa_n = sum(oa == 1, na.rm = TRUE),
    fund_n = sum(fund == 1, na.rm = TRUE),
    q1_n = sum(quartile == "Q1", na.rm = TRUE),
    q2_n = sum(quartile == "Q2", na.rm = TRUE),
    q3_n = sum(quartile == "Q3", na.rm = TRUE),
    q4_n = sum(quartile == "Q4", na.rm = TRUE),
    q_na_n = sum(is.na(quartile)),
    LS_n = sum(LS == 1, na.rm = TRUE),
    SS_n = sum(SS == 1, na.rm = TRUE),
    PS_n = sum(PS == 1, na.rm = TRUE),
    HS_n = sum(HS == 1, na.rm = TRUE),
  .groups = "drop"
  )

make_row <- function(metric, values) {
  tibble(metric = metric) |>
  bind_cols(as_tibble_row(setNames(values, mode_stats$mode)))
}

tbl1 <- bind_rows(
  make_row("N papers", mode_stats$n |> scales::comma()),
  make_row("median n_authors (IQR)", mode_stats$n_auth_iqr),
  make_row("median n_countries (IQR)", mode_stats$n_ctry_iqr),
  make_row("OA n (%)", map2_chr(mode_stats$oa_n, mode_stats$n, fmt_n_pct)),
  make_row("funded n (%)", map2_chr(mode_stats$fund_n, mode_stats$n, fmt_n_pct)),
  make_row("Q1 n (% within mode)", map2_chr(mode_stats$q1_n, mode_stats$n, fmt_n_pct)),
  make_row("Q2 n (% within mode)", map2_chr(mode_stats$q2_n, mode_stats$n, fmt_n_pct)),
  make_row("Q3 n (% within mode)", map2_chr(mode_stats$q3_n, mode_stats$n, fmt_n_pct)),
  make_row("Q4 n (% within mode)", map2_chr(mode_stats$q4_n, mode_stats$n, fmt_n_pct)),
  make_row("Quartile missing n (% within mode)", map2_chr(mode_stats$q_na_n, mode_stats$n, fmt_n_pct)),
  make_row("LS n (% within mode)", map2_chr(mode_stats$LS_n, mode_stats$n, fmt_n_pct)),
  make_row("SS n (% within mode)", map2_chr(mode_stats$SS_n, mode_stats$n, fmt_n_pct)),
  make_row("PS n (% within mode)", map2_chr(mode_stats$PS_n, mode_stats$n, fmt_n_pct)),
  make_row("HS n (% within mode)", map2_chr(mode_stats$HS_n, mode_stats$n, fmt_n_pct))
  ) |>
  select(metric, Bilateral, Multilateral, Hyperauthored, Total)

tbl1 |> knitr::kable(align = c("l", "r", "r", "r", "r"))
```

# RQ1 — Trajectory

```{r}
#| label: fig-volume

d <- dat |> filter(year >= 2000)

vol_year <- d |>
  count(year, name = "n_total") |>
  arrange(year) |>
  mutate(growth = n_total / lag(n_total) - 1)

mode_year <- d |>
  count(year, coop, name = "n_mode") |>
  left_join(vol_year, by = "year")

min_g <- suppressWarnings(min(vol_year$growth, na.rm = TRUE))
max_g <- suppressWarnings(max(vol_year$growth, na.rm = TRUE))
max_n <- max(vol_year$n_total, na.rm = TRUE)

vol_year <- vol_year |>
  mutate(growth_scaled = (growth - min_g) / (max_g - min_g) * max_n)

ggplot() +
  geom_col(
    data = mode_year, 
    aes(x = year, y = n_mode, fill = coop)
  ) +
  geom_line(
    data = vol_year,
    aes(x = year, y = growth_scaled, group = 1),
    linewidth = 0.5
  ) +
  geom_vline(xintercept = 2014, linetype = "dotted") +
  geom_vline(xintercept = 2023, linetype = "dotted") +
  scale_y_continuous(
    name = "Publications (count)",
    sec.axis = sec_axis(
      ~ . / max_n * (max_g - min_g) + min_g,
      name = "Annual growth rate",
      labels = scales::percent_format(accuracy = 1)
    )
  ) +
  labs(
    x = NULL, 
    fill = "Collaboration mode"
  )
```


```{r}
#| label: fig-fields

dat |>
  filter(year >= 2000) |>
  group_by(year) |>
  summarise(
    LS = mean(LS == 1, na.rm = TRUE),
    SS = mean(SS == 1, na.rm = TRUE),
    PS = mean(PS == 1, na.rm = TRUE),
    HS = mean(HS == 1, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_longer(
    cols = c(LS, SS, PS, HS), 
    names_to = "field", 
    values_to = "share"
  ) |>
  mutate(field = factor(field, levels = c("LS", "SS", "PS", "HS"))) |>
  ggplot(aes(x = year, y = share, color = field)) +
  geom_line() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = NULL, 
    y = "Share of publications", 
    color = NULL
  )
```


```{r}
#| label: fig-fields-stacked-normalized

dat |>
  filter(year >= 2000) |>
  group_by(year) |>
  summarise(
    LS = sum(LS == 1, na.rm = TRUE),
    SS = sum(SS == 1, na.rm = TRUE),
    PS = sum(PS == 1, na.rm = TRUE),
    HS = sum(HS == 1, na.rm = TRUE),
    .groups = "drop"
  ) |>
  # Normalize so the sum of LS, SS, PS, HS for each year equals 1
  mutate(total_hits = LS + SS + PS + HS) |>
  mutate(across(c(LS, SS, PS, HS), ~ . / total_hits)) |>
  select(-total_hits) |>
  pivot_longer(
    cols = c(LS, SS, PS, HS), 
    names_to = "field", 
    values_to = "share"
  ) |>
  mutate(field = factor(field, levels = c("LS", "PS", "HS", "SS"))) |>
  ggplot(aes(x = year, y = share, fill = field)) +
  geom_area(alpha = 0.8) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1), 
    expand = c(0, 0)
  ) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    x = NULL, 
    y = "Relative Research Effort", 
    fill = "Field"
  )
```


```{r}
#| label: fig-quality

dat |>
  filter(year >= 2000, !is.na(quartile)) |>
  count(year, quartile, name = "n") |>
  group_by(year) |>
  mutate(share = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(x = year, y = share, fill = quartile)) +
  geom_area(alpha = 0.7) +
  coord_cartesian(ylim = c(0, 1)) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1), 
    breaks = seq(0, 1, 0.2),
    expand = c(0, 0)
  ) +
  labs(
    x = NULL, 
    y = "Share of publications", 
    fill = NULL
  )
```


```{r}
#| label: fig-oa

dat |>
  filter(year >= 2000) |>
  mutate(
    oa_status = factor(if_else(oa == 1, "OA", "Non-OA"), levels = c("OA", "Non-OA"))
  ) |>
  count(year, oa_status, name = "n") |>
  group_by(year) |>
  mutate(share = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(x = year, y = share, fill = oa_status)) +
  geom_area(alpha = 0.85) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "black") +
  coord_cartesian(ylim = c(0, 1)) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1), 
    breaks = seq(0, 1, 0.2),
    expand = c(0, 0)
  ) +
  labs(
    x = NULL, 
    y = "Share of publications", 
    fill = NULL
  )
```

# RQ2


```{r}
#| label: fig-equity

dat <- dat |> mutate(parity_gap = abs(fc_vn - fc_jp))

dat |>
  filter(year >= 2000) |>
  group_by(year) |>
  summarise(
    fc_vn = mean(fc_vn, na.rm = TRUE),
    fc_jp = mean(fc_jp, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_longer(
    cols = c(fc_vn, fc_jp), 
    names_to = "series", 
    values_to = "value"
  ) |>
  mutate(
    series = recode(
      series,
      fc_vn = "Vietnam (fractional)",
      fc_jp = "Japan (fractional)"
    ),
    series = factor(
      series, 
      levels = c("Vietnam (fractional)", "Japan (fractional)")
    )
  ) |>
  {
    \(data) {
      lim <- range(data$value, na.rm = TRUE)
      pad <- max(0.02, 0.08 * diff(lim))
      
      ggplot(data, aes(x = year, y = value, color = series)) +
        geom_line() +
        coord_cartesian(ylim = c(lim[1] - pad, lim[2] + pad)) +
        scale_y_continuous(labels = percent_format(accuracy = 1)) +
        # MODIFIED: Use manual colors to match Figure 2
        scale_color_manual(values = c(
          "Vietnam (fractional)" = "#1B9E77", # Green
          "Japan (fractional)" = "#D95F02"    # Orange
        )) +
        labs(
          x = NULL, 
          y = "Mean fractional share", 
          color = NULL
        )
    }
  }()
```


```{r}
#| label: fig-leadership

lead_year <- bind_rows(
  dat |>
    filter(year >= 2000) |>
    group_by(year) |>
    summarise(value = mean(fa_vn, na.rm = TRUE), .groups = "drop") |>
    mutate(series = "Vietnam — First author"),
  dat |>
    filter(year >= 2000) |>
    group_by(year) |>
    summarise(value = mean(fa_jp, na.rm = TRUE), .groups = "drop") |>
    mutate(series = "Japan — First author"),
  dat |>
    filter(year >= 2000, !is.na(ca_vn)) |>
    group_by(year) |>
    summarise(value = mean(ca_vn, na.rm = TRUE), .groups = "drop") |>
    mutate(series = "Vietnam — Corresponding author"),
  dat |>
    filter(year >= 2000, !is.na(ca_jp)) |>
    group_by(year) |>
    summarise(value = mean(ca_jp, na.rm = TRUE), .groups = "drop") |>
    mutate(series = "Japan — Corresponding author")
) |>
  mutate(
    series = factor(
      series,
      levels = c(
        "Vietnam — First author",
        "Japan — First author",
        "Vietnam — Corresponding author",
        "Japan — Corresponding author"
      )
    )
  )

lead_year |>
  {
    \(data) {
      lim <- range(data$value, na.rm = TRUE)
      pad <- max(0.02, 0.08 * diff(lim))
      
      ggplot(data, aes(x = year, y = value, color = series, linetype = series)) +
        geom_line() +
        coord_cartesian(ylim = c(lim[1] - pad, lim[2] + pad)) +
        scale_y_continuous(labels = percent_format(accuracy = 1)) +
        # MODIFIED: Assign same color to same country, regardless of author type
        scale_color_manual(values = c(
          "Vietnam — First author" = "#1B9E77",         # Green
          "Vietnam — Corresponding author" = "#1B9E77", # Green
          "Japan — First author" = "#D95F02",           # Orange
          "Japan — Corresponding author" = "#D95F02"    # Orange
        )) +
        # Keep linetypes to distinguish author role
        scale_linetype_manual(values = c("dashed", "dashed", "solid", "solid")) +
        labs(
          x = NULL, 
          y = "Leadership share", 
          color = NULL, 
          linetype = NULL
        )
    }
  }()
```

# RQ3

This chunk constructs a time- and field-aware citation impact normalization that preserves the multi-label nature of Scopus macro-field tagging (HS, LS, SS, PS). Because many papers are assigned to multiple macro fields, forcing a single-field classification would be arbitrary and could distort comparisons. We therefore compute year-specific field baselines and then normalize each paper relative to the expected citation level implied by its own field portfolio.

Concretely, we first reshape the dataset into a long format to estimate, for each year × macro field, the average citation count (`mu_cited`) and the average log-citation (`mu_log1p = mean(log1p(cited))`). Next, for each paper, we calculate the number of macro fields it is tagged in (`k_field`) and assign equal fractional weights (1/`k_field`) across its tagged fields. Using these weights, we derive an expected baseline for each paper as the weighted average of the relevant year × field baselines. Finally, we construct two normalized impact measures: (i) a ratio-based index, `impact_norm = cited / impact_mu`, and (ii) a log-centered index, `impact_norm_log = log1p(cited) − impact_mu_log`. The log-centered form is typically more stable in citation data because it reduces sensitivity to extreme outliers and heavy right tails, while still retaining the interpretation of relative performance against a year- and field-adjusted benchmark.

Let paper (i) be published in year (y_i). Let the set of macro fields be (F={\text{HS},\text{LS},\text{SS},\text{PS}}). Define an indicator (I_{if}\in{0,1}) that paper (i) is tagged in field (f).

### Step 1. Year–field baselines

[
\mu^{(c)}*{f,y}=\mathbb{E}!\left[C \mid f,y\right]
]
[
\mu^{(\ell)}*{f,y}=\mathbb{E}!\left[\log(1+C)\mid f,y\right]
]
where (C) is the citation count.

### Step 2. Fractional weights for multi-field papers

[
k_i=\sum_{f\in F} I_{if}
\qquad\text{and}\qquad
w_{if}=\frac{I_{if}}{k_i}
]

### Step 3. Paper-specific expected baselines (mixture over fields)

[
\mu^{(c)}*{i}=\sum*{f\in F} w_{if},\mu^{(c)}*{f,y_i}
]
[
\mu^{(\ell)}*{i}=\sum_{f\in F} w_{if},\mu^{(\ell)}_{f,y_i}
]

### Step 4. Normalized impact measures

**Ratio normalization**
[
\text{impact_norm}*i=\frac{C_i}{\mu^{(c)}*{i}}
]

**Log-centered normalization**
[
\text{impact_norm_log}*i=\log(1+C_i)-\mu^{(\ell)}*{i}
]

These are exactly the quantities computed in the chunk: the paper’s expected baseline is the weighted average of year×field means, with equal fractional weight across all macro fields assigned to that paper.


```{r}
#| label: impact-norm-func

macro_fields <- c("HS", "LS", "SS", "PS")

make_impact_norm <- function(df) {
  field_year_mu <- df |>
    pivot_longer(all_of(macro_fields), names_to = "field", values_to = "flag") |>
    filter(flag == 1) |>
    group_by(year, field) |>
    summarise(
      mu_cited = mean(cited, na.rm = TRUE),
      mu_log1p = mean(log1p(cited), na.rm = TRUE),
      .groups = "drop"
    )
  
  df |>
    mutate(k_field = rowSums(across(all_of(macro_fields)), na.rm = TRUE)) |>
    pivot_longer(all_of(macro_fields), names_to = "field", values_to = "flag") |>
    filter(flag == 1, k_field > 0) |>
    left_join(field_year_mu, by = c("year", "field")) |>
    mutate(w = 1 / k_field) |>
    group_by(id) |>
    summarise(
      impact_mu = sum(w * mu_cited, na.rm = TRUE),
      impact_mu_log = sum(w * mu_log1p, na.rm = TRUE),
      .groups = "drop"
    ) |>
    right_join(df, by = "id") |>
    mutate(
      impact_mu = if_else(is.na(impact_mu) | impact_mu <= 0, NA_real_, impact_mu),
      impact_norm = cited / impact_mu,
      impact_norm_log = log1p(cited) - impact_mu_log
    )
}
```


```{r}
#| label: apply-norm

dat <- make_impact_norm(dat)
dat_core <- make_impact_norm(dat_core)
dat_bi <- make_impact_norm(dat_bi)
```


```{r}
#| label: fig-impact-parity

d <- dat_core |>
  filter(year >= 2000, !is.na(ca_vn), !is.na(ca_jp)) |>
  mutate(
    lead = case_when(
      ca_vn > 0 & ca_jp == 0 ~ "Vietnam-led (CA)",
      ca_jp > 0 & ca_vn == 0 ~ "Japan-led (CA)",
      TRUE ~ NA_character_
    )
  ) |>
  filter(!is.na(lead)) |>
  mutate(lead = factor(lead, levels = c("Vietnam-led (CA)", "Japan-led (CA)")))

parity_year <- d |>
  group_by(year, lead) |>
  summarise(impact = mean(impact_norm_log, na.rm = TRUE), .groups = "drop") |>
  arrange(year)

gap_year <- parity_year |>
  select(year, lead, impact) |>
  pivot_wider(names_from = lead, values_from = impact) |>
  transmute(year, gap = `Vietnam-led (CA)` - `Japan-led (CA)`)

p <- ggplot(parity_year, aes(year, impact, color = lead)) +
  geom_line(linewidth = 0.9) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = NULL, y = "Mean normalized impact (log-centered)", color = NULL)

p +
  geom_line(
    data = gap_year,
    aes(year, gap),
    inherit.aes = FALSE,
    linewidth = 0.8,
    linetype = "dashed",
    color = "grey30"
  ) +
  labs(y = "Mean normalized impact (log-centered); dashed = VN − JP gap")
```


```{r}
#| label: model-impact

d <- dat_core |>
  filter(year >= 2000, !is.na(ca_vn), !is.na(ca_jp), !is.na(impact_norm_log)) |>
  mutate(
    lead_vn = as.integer(ca_vn > 0 & ca_jp == 0),
    lead_jp = as.integer(ca_jp > 0 & ca_vn == 0),
    lead = case_when(lead_vn == 1 ~ "VN", lead_jp == 1 ~ "JP", TRUE ~ NA_character_)
  ) |>
  filter(!is.na(lead)) |>
  mutate(
    lead = factor(lead, levels = c("JP", "VN")),
    quartile = fct_drop(quartile)
  )

m0 <- lm(impact_norm_log ~ lead * year, data = d)

m1 <- lm(impact_norm_log ~ lead * year + HS + LS + SS + PS, data = d)

m2 <- lm(
  impact_norm_log ~ lead * year + HS + LS + SS + PS +
    quartile + oa + fund + log1p(n_authors) + n_countries,
  data = d
)

broom::tidy(m2) |>
  filter(term %in% c("leadVN", "year", "leadVN:year")) |>
  knitr::kable(digits = 3)
```


```{r}
#| label: predict-impact-gap

yrs <- sort(unique(d$year))

pred <- tidyr::expand_grid(
  year = yrs,
  lead = levels(d$lead)
) |>
  mutate(
    HS = 0L, LS = 0L, SS = 0L, PS = 0L,
    quartile = levels(d$quartile)[1],
    oa = 0L, fund = 0L,
    n_authors = median(d$n_authors, na.rm = TRUE),
    n_countries = median(d$n_countries, na.rm = TRUE)
  ) |>
  mutate(
    lead = factor(lead, levels = levels(d$lead)),
    quartile = factor(quartile, levels = levels(d$quartile), 
                      ordered = is.ordered(d$quartile))
  )

pp <- predict(m2, newdata = pred, se.fit = TRUE)

pred <- pred |>
  mutate(fit = as.numeric(pp$fit), se = as.numeric(pp$se.fit))

diff_year <- pred |>
  select(year, lead, fit, se) |>
  tidyr::pivot_wider(names_from = lead, values_from = c(fit, se)) |>
  transmute(
    year,
    diff = .data$fit_VN - .data$fit_JP,
    se   = sqrt(.data$se_VN^2 + .data$se_JP^2),
    lo   = diff - 1.96 * se,
    hi   = diff + 1.96 * se
  )

ggplot(diff_year, aes(year, diff)) +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.2) +
  geom_line() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = NULL, y = "Estimated VN − JP difference (normalized impact, log-centered)")
```


```{r}
#| label: robustness-fa

d_fa <- dat_core |>
  filter(year >= 2000, !is.na(impact_norm_log)) |>
  mutate(
    lead_vn = as.integer(fa_vn > 0 & fa_jp == 0),
    lead_jp = as.integer(fa_jp > 0 & fa_vn == 0),
    lead = case_when(lead_vn == 1 ~ "VN", lead_jp == 1 ~ "JP", TRUE ~ NA_character_)
  ) |>
  filter(!is.na(lead)) |>
  mutate(lead = factor(lead, levels = c("JP", "VN")))

m2_fa <- lm(
  impact_norm_log ~ lead * year + HS + LS + SS + PS +
    quartile + oa + fund + log1p(n_authors) + n_countries,
  data = d_fa
)

broom::tidy(m2_fa) |>
  filter(term %in% c("leadVN", "year", "leadVN:year")) |>
  knitr::kable(digits = 3)
```


```{r}
#| label: robustness-bi

d_bi <- dat_bi |>
  filter(year >= 2000, !is.na(ca_vn), !is.na(ca_jp), !is.na(impact_norm_log)) |>
  mutate(
    lead_vn = as.integer(ca_vn > 0 & ca_jp == 0),
    lead_jp = as.integer(ca_jp > 0 & ca_vn == 0),
    lead = case_when(lead_vn == 1 ~ "VN", lead_jp == 1 ~ "JP", TRUE ~ NA_character_)
  ) |>
  filter(!is.na(lead)) |>
  mutate(lead = factor(lead, levels = c("JP", "VN")))

m2_bi <- lm(
  impact_norm_log ~ lead * year + HS + LS + SS + PS +
    quartile + oa + fund + log1p(n_authors) + n_countries,
  data = d_bi
)

broom::tidy(m2_bi) |>
  filter(term %in% c("leadVN", "year", "leadVN:year")) |>
  knitr::kable(digits = 3)
```


```{r}
#| label: predict-impact-gap-bi

# Get all unique years in bilateral data (using d_bi, not dat_bi)
yrs_bi <- sort(unique(d_bi$year))

# Create prediction grid for bilateral subset
pred_bi <- tidyr::expand_grid(
  year = yrs_bi,
  lead = levels(d_bi$lead)
) |>
  mutate(
    HS = 0L, LS = 0L, SS = 0L, PS = 0L,
    quartile = levels(d_bi$quartile)[1],
    oa = 0L, fund = 0L,
    n_authors = median(d_bi$n_authors, na.rm = TRUE),
    n_countries = median(d_bi$n_countries, na.rm = TRUE)
  ) |>
  mutate(
    lead = factor(lead, levels = levels(d_bi$lead)),
    # Corrected the typo here: levels(d_bi$quartile) instead of dat_bi(...)
    quartile = factor(quartile, levels = levels(d_bi$quartile), 
                      ordered = is.ordered(d_bi$quartile))
  )

# Generate predictions
pp_bi <- predict(m2_bi, newdata = pred_bi, se.fit = TRUE)

pred_bi <- pred_bi |>
  mutate(fit = as.numeric(pp_bi$fit), se = as.numeric(pp_bi$se.fit))

# Calculate differences
diff_year_bi <- pred_bi |>
  select(year, lead, fit, se) |>
  tidyr::pivot_wider(names_from = lead, values_from = c(fit, se)) |>
  transmute(
    year,
    diff = .data$fit_VN - .data$fit_JP,
    se   = sqrt(.data$se_VN^2 + .data$se_JP^2),
    lo   = diff - 1.96 * se,
    hi   = diff + 1.96 * se
  )

# Plot bilateral-only predictions
ggplot(diff_year_bi, aes(year, diff)) +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.2) +
  geom_line() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = NULL, y = "Bilateral-only: Estimated VN − JP difference (normalized impact, log-centered)")
```


```{r}
library(tidyverse)
library(broom)

tidy(m0) |>
  filter(term %in% c("leadVN", "year", "leadVN:year")) |>
  select(term, estimate, std.error, statistic, p.value) |>
  print()

tidy(m1) |>
  filter(term %in% c("leadVN", "year", "leadVN:year")) |>
  select(term, estimate, std.error, statistic, p.value) |>
  print()

tidy(m2) |>
  filter(term %in% c("leadVN", "year", "leadVN:year")) |>
  select(term, estimate, std.error, statistic, p.value) |>
  print()

model_comparison <- tibble(
  Model = c("M0: Baseline", "M1: + Fields", "M2: Full"),
  R_squared = c(
    summary(m0)$r.squared,
    summary(m1)$r.squared,
    summary(m2)$r.squared
  ),
  Adj_R_squared = c(
    summary(m0)$adj.r.squared,
    summary(m1)$adj.r.squared,
    summary(m2)$adj.r.squared
  ),
  N_obs = c(
    nobs(m0),
    nobs(m1),
    nobs(m2)
  )
)

print(model_comparison)

tidy(m2_fa) |>
  filter(term %in% c("leadVN", "year", "leadVN:year")) |>
  select(term, estimate, std.error, statistic, p.value) |>
  print()

tidy(m2_bi) |>
  filter(term %in% c("leadVN", "year", "leadVN:year")) |>
  select(term, estimate, std.error, statistic, p.value) |>
  print()

gap_summary <- diff_year |>
  summarise(
    Min_gap = min(diff),
    Max_gap = max(diff),
    Mean_gap = mean(diff),
    Gap_2000 = diff[year == 2000],
    Gap_2023 = diff[year == max(year)],
    Change = Gap_2023 - Gap_2000
  )

print(gap_summary)

zero_crossing <- diff_year |>
  mutate(sign_change = sign(diff) != sign(lag(diff))) |>
  filter(sign_change == TRUE) |>
  select(year, diff, lo, hi)

if (nrow(zero_crossing) > 0) {
  print(zero_crossing)
}

parity_stats <- d |>
  group_by(lead) |>
  summarise(
    N = n(),
    Mean_impact = mean(impact_norm_log, na.rm = TRUE),
    Median_impact = median(impact_norm_log, na.rm = TRUE),
    SD_impact = sd(impact_norm_log, na.rm = TRUE),
    Q25 = quantile(impact_norm_log, 0.25, na.rm = TRUE),
    Q75 = quantile(impact_norm_log, 0.75, na.rm = TRUE)
  )

print(parity_stats)

temporal_comparison <- d |>
  mutate(period = case_when(
    year >= 2000 & year <= 2010 ~ "Early (2000-2010)",
    year >= 2014 & year <= 2023 ~ "Recent (2014-2023)",
    TRUE ~ NA_character_
  )) |>
  filter(!is.na(period)) |>
  group_by(period, lead) |>
  summarise(
    N = n(),
    Mean_impact = mean(impact_norm_log, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_wider(names_from = lead, values_from = c(N, Mean_impact))

print(temporal_comparison)

if (nrow(temporal_comparison) == 2) {
  early_gap <- temporal_comparison$`Mean_impact_Vietnam-led (CA)`[1] - 
               temporal_comparison$`Mean_impact_Japan-led (CA)`[1]
  recent_gap <- temporal_comparison$`Mean_impact_Vietnam-led (CA)`[2] - 
                temporal_comparison$`Mean_impact_Japan-led (CA)`[2]
  
  cat(sprintf("Early period gap: %.4f\n", early_gap))
  cat(sprintf("Recent period gap: %.4f\n", recent_gap))
  cat(sprintf("Change: %.4f (%+.1f%%)\n", 
              recent_gap - early_gap, 
              100 * (recent_gap - early_gap) / abs(early_gap)))
}

interaction_coef <- coef(m2)["leadVN:year"]
interaction_pval <- summary(m2)$coefficients["leadVN:year", "Pr(>|t|)"]

cat(sprintf("Estimate: %.6f\n", interaction_coef))
cat(sprintf("P-value: %.4f\n", interaction_pval))

if (interaction_pval < 0.001) {
  sig_level <- "***"
} else if (interaction_pval < 0.01) {
  sig_level <- "**"
} else if (interaction_pval < 0.05) {
  sig_level <- "*"
} else if (interaction_pval < 0.10) {
  sig_level <- "†"
} else {
  sig_level <- "n.s."
}

if (interaction_coef > 0 & interaction_pval < 0.05) {
  cat("The Vietnam-Japan impact gap is NARROWING over time ", sig_level, "\n")
} else if (interaction_coef < 0 & interaction_pval < 0.05) {
  cat("The Vietnam-Japan impact gap is WIDENING over time ", sig_level, "\n")
} else {
  cat("No significant temporal trend in the impact gap ", sig_level, "\n")
}
```

# Diagnostic

```{r}
#| label: diagnostics-vif

# Load car package for VIF (install if needed: install.packages("car"))
if (!require(car)) install.packages("car")
library(car)

cat("=== MULTICOLLINEARITY CHECK (VIF) ===\n")
cat("Note: VIF > 5 or 10 indicates problematic collinearity.\n\n")

# Check VIF on Model 2 (Full Model)
# We must use type="predictor" or construct a model without interaction 
# to avoid artificial inflation from the product term (lead * year)
m2_no_int <- lm(
  impact_norm_log ~ lead + year + HS + LS + SS + PS +
    quartile + oa + fund + log1p(n_authors) + n_countries,
  data = d
)

vif_vals <- car::vif(m2_no_int)
print(vif_vals)

# Visual check: If all values are < 5, you are safe.
if(any(vif_vals > 5)) {
  cat("\nWARNING: Some variables have high VIF (>5).\n")
} else {
  cat("\nSUCCESS: No severe multicollinearity detected among controls.\n")
}
```


```{r}
#| label: diagnostics-plots

# Set up 2x2 plotting area
par(mfrow = c(2, 2))

# Plot diagnostics for the full model (m2)
plot(m2, which = c(1, 2, 3, 5))

# Reset plotting area
par(mfrow = c(1, 1))

# --- PLOT INTERPRETATION ---
# 1. Residuals vs Fitted (Top-Left): Should look like a random cloud (linearity).
# 2. Normal Q-Q (Top-Right): Dots should hug the diagonal line (normality).
# 3. Scale-Location (Bottom-Left): Line should be horizontal (homoscedasticity).
# 4. Residuals vs Leverage (Bottom-Right): Look for points outside dashed Cook's distance lines.
```


```{r}
#| label: diagnostics-heteroscedasticity

library(lmtest) # For bptest

cat("=== BREUSCH-PAGAN TEST FOR HETEROSCEDASTICITY ===\n")
cat("Null Hypothesis: Variance is constant (Homoscedasticity)\n")
cat("P-value < 0.05 implies Heteroscedasticity (bad)\n\n")

bp_test <- lmtest::bptest(m2)
print(bp_test)

if(bp_test$p.value < 0.05) {
  cat("\nRESULT: Heteroscedasticity detected.\n")
  cat("ACTION: You should report 'Robust Standard Errors' (Step 4).\n")
} else {
  cat("\nRESULT: Homoscedasticity holds. Standard OLS errors are valid.\n")
}
```


```{r}
#| label: robust-se

# Load sandwich package for robust covariance matrix
if (!require(sandwich)) install.packages("sandwich")
library(sandwich)
library(lmtest)

cat("=== ROBUST STANDARD ERRORS (HC1) ===\n")
cat("Re-estimating significance adjusting for heteroscedasticity...\n\n")

# Calculate Robust Coefficients
robust_m2 <- lmtest::coeftest(m2, vcov = sandwich::vcovHC(m2, type = "HC1"))

# Extract and print the key variables for RQ3
# We look at the rows corresponding to our interaction term
robust_tidy <- broom::tidy(robust_m2) |>
  filter(term %in% c("leadVN", "year", "leadVN:year"))

print(robust_tidy)

cat("\nComparison: Check if the 'p.value' for 'leadVN:year' is still significant.\n")
cat("If yes, your finding is robust to citation variance issues.")
```


```{r}
# ==============================================================================
# VARIABLE DIAGNOSTICS: IMPACT_NORM_LOG
# ==============================================================================

# 1. Descriptive Statistics Table
desc_stats <- d |>
  summarise(
    n = n(),
    mean = mean(impact_norm_log, na.rm = TRUE),
    median = median(impact_norm_log, na.rm = TRUE),
    sd = sd(impact_norm_log, na.rm = TRUE),
    min = min(impact_norm_log, na.rm = TRUE),
    max = max(impact_norm_log, na.rm = TRUE),
    skewness = (sum((impact_norm_log - mean(impact_norm_log, na.rm = TRUE))^3) / n()) / 
               (sd(impact_norm_log, na.rm = TRUE)^3)
  )

print(desc_stats)

# 2. Diagnostic Histogram
# Purpose: Check if the distribution is roughly bell-shaped (ideal for OLS)
# and centered near 0 (the expected baseline).

ggplot(d, aes(x = impact_norm_log)) +
  # Histogram with reasonable bin width
  geom_histogram(aes(y = after_stat(density)), bins = 60, 
                 fill = "#69b3a2", color = "white", alpha = 0.8) +
  # Overlay normal curve for comparison
  stat_function(fun = dnorm, 
                args = list(mean = mean(d$impact_norm_log, na.rm = TRUE), 
                            sd = sd(d$impact_norm_log, na.rm = TRUE)),
                color = "darkred", linewidth = 1) +
  # Vertical line at 0 (Impact Parity)
  geom_vline(xintercept = 0, linetype = "dashed", size = 0.8) +
  labs(
    title = "Distribution of Log-Normalized Impact",
    subtitle = "Red curve = Normal distribution; Dashed line = Field average (0)",
    x = "Normalized Impact (Log-Centered)",
    y = "Density"
  ) +
  theme_minimal()
```


```{r}

```