---
title: "Data Processing Pipeline"
format: html
execute:
    echo: true
    warning: true
    message: true
---

# Setting Up

This initial section establishes the computational environment by loading essential R libraries: `tidyverse` for general data manipulation and visualization, `countrycode` for standardizing geographic names to facilitate accurate collaboration mapping, and `text2sdg` for extracting Sustainable Development Goals from text. 

We define the directory paths for raw data and funding inputs to ensure reproducible relative paths. The raw Scopus dataset (`all_1227.csv`) is imported, after which we filter for only the essential columns required for bibliometric analysis to optimize memory usage. Variable names are standardized from the raw Scopus format to snake_case for coding consistency. Finally, three key preliminary variables are constructed: `id` assigns a sequential row number as a unique identifier for this processing session; `oa` transforms the textual 'Open Access' status into a clean binary integer (1 for Open Access, 0 otherwise); and `n_authors` is calculated by counting semicolon separators in the `authors_str` column, serving as a baseline integer count of contributors.


```{r}
#| label: setup

library(tidyverse)
library(countrycode)
library(text2sdg)

DATA_DIR <- "data"
FUNDER_DIR <- file.path(DATA_DIR, "funder_data")

dat_reg <- read_csv(file.path(DATA_DIR, "all_1227.csv"), show_col_types = FALSE) |>
  select(EID, Year, "Source title", ISSN, "Cited by", Authors, "Authors with affiliations", DOI, "Open Access", Abstract, "Correspondence Address", "Author Keywords") |>
  rename(
    eid = EID,
    year = Year,
    title = "Source title",
    issn = ISSN,
    cited = "Cited by",
    authors_str = Authors,
    auth_aff = "Authors with affiliations",
    doi = DOI,
    open_access = "Open Access",
    abstract = Abstract,
    corr = "Correspondence Address",
    auth_kw = "Author Keywords"
  ) |>
  mutate(
    id = row_number(),
    oa = if_else(!is.na(open_access) & open_access != "", 1L, 0L),
    n_authors = str_count(authors_str, ";") + 1L
  ) |>
  select(-open_access)
```

# Journal-level data

This section enriches the bibliographic dataset by integrating external journal-level metadata, which is crucial for assessing the disciplinary composition and quality of research outlets. We perform two sequential left_join operations using the journal title as the key. First, we merge data from `sources_1227.csv` to map each publication to Scopus Subject Areas (e.g., Engineering, Medicine, Social Sciences). These are aggregated into specific binary flags for 27 distinct disciplines and 4 broader Macro Areas (Life Sciences, Social Sciences, Physical Sciences, and Health Sciences). Second, we incorporate quality metrics from `sjr_1227.csv` (SCImago Journal Rank), extracting the raw sjr_score as a continuous measure of journal prestige and the quartile ranking (Q1 to Q4) to categorize journals based on their impact factor distribution within their field. This step ensures that every paper is contextualized within its specific scientific domain and publication tier.


```{r}
#| label: process-journal

dat_reg <- dat_reg |>
  left_join(
    read_csv(file.path(DATA_DIR, "sources_1227.csv"), show_col_types = FALSE) |>
      transmute(
        title = `Source Title`,
        LS = `Top level:\n\nLife Sciences`,
        SS = `Top level:\n\nSocial Sciences`,
        PS = `Top level:\n\nPhysical Sciences`,
        HS = `Top level:\n\nHealth Sciences`,
        mult     = `1000 \nGeneral`,
        agr_bio  = `1100\nAgricultural and Biological Sciences`,
        arts     = `1200\nArts and Humanities`,
        bioch    = `1300\nBiochemistry, Genetics and Molecular Biology`,
        bus      = `1400\nBusiness, Management and Accounting`,
        chem_eng = `1500\nChemical Engineering`,
        chem     = `1600\nChemistry`,
        comp     = `1700\nComputer Science`,
        deci     = `1800\nDecision Sciences`,
        earth    = `1900\nEarth and Planetary Sciences`,
        econ     = `2000\nEconomics, Econometrics and Finance`,
        ener     = `2100\nEnergy`,
        engi     = `2200\nEngineering`,
        env      = `2300\nEnvironmental Science`,
        immu     = `2400\nImmunology and Microbiology`,
        mate     = `2500\nMaterials Science`,
        math     = `2600\nMathematics`,
        medi     = `2700\nMedicine`,
        neur     = `2800\nNeuroscience`,
        nurs     = `2900\nNursing`,
        phar     = `3000\nPharmacology, Toxicology and Pharmaceutics`,
        phys     = `3100\nPhysics and Astronomy`,
        psyc     = `3200\nPsychology`,
        soci     = `3300\nSocial Sciences`,
        vete     = `3400\nVeterinary`,
        dent     = `3500\nDentistry`,
        heal     = `3600\nHealth Professions`
      ) |>
      mutate(across(-title, ~ if_else(!is.na(.), 1L, 0L))) |>
      distinct(title, .keep_all = TRUE),
    by = "title"
  ) |>
  left_join(
    read_delim(file.path(DATA_DIR, "sjr_1227.csv"), delim = ";", show_col_types = FALSE) |>
      filter(Type == "journal") |>
      transmute(
        title = Title,
        sjr_score = parse_number(SJR, locale = locale(grouping_mark = ",")) / 1000,
        quartile = factor(`SJR Best Quartile`, levels = c("Q1", "Q2", "Q3", "Q4"))
      ) |>
      distinct(title, .keep_all = TRUE),
    by = "title"
  )
```

# Collaboration

This section establishes the geographic scope of each publication by detecting and unifying country affiliations. We define a comprehensive set of search terms covering common abbreviations (e.g., "USA", "PRC", "UAE") and regional components (e.g., "England", "Scotland", "Wales"). To ensure accurate counting, the process involves two steps: first, extracting all potential country mentions using a strictly bounded regular expression; second, passing these tokens through countrycode to unify them into standardized country names (e.g., "England", "UK", and "Great Britain" are all mapped to a single entity: "United Kingdom"). We then calculate n_countries based on these unique standardized entities to classify papers as "Domestic" (1), "Bilateral" (2), or "Multilateral" (>2), while separately flagging "Hyperauthored" papers (≥100 authors).

```{r}
#| label: process-collaboration

country_list <- c(
  countrycode::codelist$country.name.en, 
  # Asia
  "Viet Nam", "Vietnam", 
  "South Korea", "Republic of Korea", "Korea", 
  "China", "People's Republic of China", "PRC",
  # Europe
  "Russia", "Russian Federation", 
  "UK", "United Kingdom",
  "Italy", "Italia",
  # Americas
  "USA", "United States", "US", "United States of America", "U.S.A.",
  # Middle East
  "UAE", "United Arab Emirates"
)

country_regex <- paste0("\\b(", paste(unique(country_list), collapse = "|"), ")\\b")

dat_reg <- dat_reg |>
  mutate(
    countries = map(auth_aff, \(x) {
      str_extract_all(x, regex(country_regex, ignore_case = TRUE))[[1]] |>
        countrycode("country.name", "country.name.en", warn = FALSE) |>
        na.omit() |>
        unique()
    }),
    n_countries = map_int(countries, length),
    coop = case_when(
      n_authors >= 100 ~ "Hyperauthored",
      n_countries == 2 ~ "Bilateral",
      n_countries > 2 ~ "Multilateral"
    )
  ) |>
  select(-countries)
```


# Authorship

4. Authorship and Leadership
This section generates granular authorship and leadership metrics. We iterate through each individual author in the auth_aff string to determine their country affiliations (Japan, Vietnam, or Other). Based on these individual attributes, we calculate:

- Full Counts (`full_`*): The distinct number of authors affiliated with each region (Authors with dual VN-JP affiliations contribute to both counts; "Other" captures authors with neither affiliation).

- Fractional Shares (`fc_`*): The proportion of total country "claims" held by each region.

- Modified Fractional Counts (`mfc_`*): The full count weighted by the inverse square root of the paper's total unique countries, a standard measure to normalize for team size and multinationality.

- Leadership (`fa_`*, `ca_`*): Binary flags indicating whether the First Author or Corresponding Author is affiliated with Vietnam, Japan, or a third country.

```{r}
library(tidyverse)

country_re <- regex(country_regex, ignore_case = TRUE)
jp_pat     <- "Japan"
vn_pat     <- "Viet\\s*Nam|Vietnam"

process_author_credits <- function(aff_string) {
  if (is.na(aff_string) || !nzchar(aff_string)) return(NULL)
  
  authors <- str_split_1(aff_string, ";") |> str_trim()
  n_authors <- length(authors)
  
  auth_data <- map_dfr(seq_along(authors), \(i) {
    countries <- str_extract_all(authors[i], country_re)[[1]] |> 
      unique()
    
    if (length(countries) == 0) countries <- "Other"
    
    is_jp <- str_detect(countries, regex(jp_pat, ignore_case = TRUE))
    is_vn <- str_detect(countries, regex(vn_pat, ignore_case = TRUE))
    is_o  <- !(is_jp | is_vn)
    
    n_c <- length(countries)
    
    tibble(
      is_fa = i == 1,
      jp_share = sum(is_jp) / n_c,
      vn_share = sum(is_vn) / n_c,
      o_share  = sum(is_o) / n_c
    )
  })

  tibble(
    full_jp = sum(auth_data$jp_share > 0),
    full_vn = sum(auth_data$vn_share > 0),
    full_o  = sum(auth_data$o_share > 0),
    fc_jp   = sum(auth_data$jp_share) / n_authors,
    fc_vn   = sum(auth_data$vn_share) / n_authors,
    fc_o    = sum(auth_data$o_share) / n_authors,
    fa_jp   = auth_data$jp_share[1],
    fa_vn   = auth_data$vn_share[1],
    fa_o    = auth_data$o_share[1]
  )
}

parse_corr_credits <- function(corr_string) {
  countries <- str_extract_all(corr_string, country_re)[[1]] |> unique()
  if (length(countries) == 0) return(tibble(ca_jp = 0, ca_vn = 0, ca_o = 1))
  
  n_c <- length(countries)
  is_jp <- str_detect(countries, regex(jp_pat, ignore_case = TRUE))
  is_vn <- str_detect(countries, regex(vn_pat, ignore_case = TRUE))
  
  tibble(
    ca_jp = sum(is_jp) / n_c,
    ca_vn = sum(is_vn) / n_c,
    ca_o  = sum(!(is_jp | is_vn)) / n_c
  )
}

dat_reg <- dat_reg |>
  left_join(
    dat_reg |> 
      distinct(auth_aff) |> 
      mutate(res = map(auth_aff, process_author_credits)) |> 
      unnest(res), 
    by = "auth_aff"
  ) |>
  left_join(
    dat_reg |> 
      distinct(corr) |> 
      mutate(res = map(corr, parse_corr_credits)) |> 
      unnest(res), 
    by = "corr"
  )

```

# Funding

This section integrates external funding data to characterize the financial support behind collaborative research. We aggregate data from eight separate CSV files, each representing a specific funding origin (Japan, Vietnam, Other) and sector (Academia, Public, Industry). These files are read and mapped to binary indicators (`fund_jp`, `fund_vn`, `fund_o` for origin; `aca`, `pub`, `ind` for sector). We consolidate these records by DOI to handle papers receiving support from multiple sources. Finally, we join this aggregated funding data with the main dataset and create a composite fund flag to identify papers with any reported funding.

```{r}
#| label: process-funding

funder_files <- tribble(
  ~file,        ~reg_name, ~sec_name,
  "ja_ind.csv", "fund_jp", "ind",
  "ja_pub.csv", "fund_jp", "pub",
  "ja_aca.csv", "fund_jp", "aca",
  "vn_pub.csv", "fund_vn", "pub",
  "vn_aca.csv", "fund_vn", "aca",
  "o_ind.csv",  "fund_o",  "ind",
  "o_pub.csv",  "fund_o",  "pub",
  "o_aca.csv",  "fund_o",  "aca"
)

# Join funding data using eid as the bridge
dat_reg <- dat_reg |>
  left_join(
    funder_files |>
      pmap_dfr(function(file, reg_name, sec_name) {
        path <- file.path(FUNDER_DIR, file)
        if (!file.exists(path)) return(NULL)
        read_csv(path, show_col_types = FALSE) |>
          rename_with(tolower) |>
          select(eid) |>
          mutate(!!reg_name := 1L, !!sec_name := 1L)
      }) |>
      group_by(eid) |>
      summarise(across(everything(), \(x) as.integer(any(x == 1, na.rm = TRUE))), .groups = "drop"),
    by = "eid"
  ) |>
  mutate(
    across(c(fund_jp, fund_vn, fund_o, ind, pub, aca), \(x) replace_na(x, 0L)),
    fund = as.integer(fund_jp | fund_vn | fund_o)
  )
```


# SDG

This final section analyzes the thematic content of the research by detecting Sustainable Development Goals (SDGs) within the abstracts using the 'Aurora' classification system via the `text2sdg` package. We transform the detection results into 17 binary columns (`sdg01` through `sdg17`), indicating the presence of each goal, and calculate a n_sdg score representing the total count of SDGs addressed per paper. Finally, we prune the dataset to retain only the analytical variables specified in the study design—bibliometric metadata, collaboration metrics, granular authorship counts (Full, Fractional, Modified Fractional), leadership flags, funding sources, and SDG indicators—and save the streamlined dataset as an RDS file for subsequent analysis.

```{r}
#| label: process-sdg-save
sdg_hits <- detect_sdg_systems(
  text = replace_na(dat_reg$abstract, ""),
  system = c("Aurora", "Elsevier", "SIRIS", "SDSN", "SDGO")
)

dat_reg <- dat_reg |>
  mutate(row_id = row_number()) |>
  left_join(
    sdg_hits |>
      mutate(
        document = as.integer(document),
        sdg_num = str_extract(sdg, "\\d+"),
        sdg_col = paste0("sdg", str_pad(sdg_num, 2, pad = "0"))
      ) |>
      distinct(document, sdg_col) |>
      pivot_wider(names_from = sdg_col, values_from = sdg_col, 
                  values_fn = \(x) 1L, values_fill = 0L),
    by = c("row_id" = "document")
  ) |>
  mutate(
    across(starts_with("sdg"), \(x) replace_na(x, 0L)),
    n_sdg = rowSums(across(starts_with("sdg")))
  ) |>
  select(
    id, year, cited, oa,
    LS, SS, PS, HS, sjr_score, quartile,
    n_authors, n_countries, coop,
    full_jp, full_vn, full_o,
    fc_jp, fc_vn, fc_o,
    fa_vn, fa_jp, fa_o,
    ca_vn, ca_jp, ca_o,
    fund_vn, fund_jp, fund_o, aca, pub, ind, fund,
    matches("sdg\\d+"), n_sdg
  ) |> 
  drop_na(LS)

write_csv(dat_reg, file.path(DATA_DIR, "updated_data.csv"))
```


```{r}

```